---
title: "Implementation details"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Implementation details}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library("ESQhtpbpk")
```


## Generic PBPK model structure

A generic PBPK model will be automatically created for each species, partition coefficient method (PC), 
cellular permeability method (CP) as well as each required combination of processes for a certain number 
of compounds used. For each generic model structure the required administrations from all studies 
using this generic model will be checked and added.

Note: CP calculation method "Charge dependent Schmitt normalized to
PK-Sim" is not supported as it is not properly implemented in the
current PK-Sim versions (11.3 and 12.0).

## Compound

The `Compound` class is used to create the compound object that will be used in the studies.
Some compound properties are already predefined and can be directly used. 

Pre-specified compound properties (with their default values) are:

```{r, eval = TRUE}
compound <- Compound$new(ID = "Acetaminophen")
compound
```

### Compound Properties

Additional compound properties can be added as needed with addProperty (with corresponding paths for 
predefined PKML files). Existing ones can be easily changed with the setPropertyValue function.
The property `name` is an alias to easily access a property in the compound object. The `parName` 
needs to correspond to the parameter name as defined in the PK-Sim. 

```{r, eval = TRUE}
compound$setPropertyValue("Molecular weight", 151.16, unit = "g/mol")

compound$addProperty(
  name = "Pint", 
  parName = "Intestinal permeability (transcellular)",
  dimension = "Velocity",
  value = 2e-5,
  unit = "cm/min"
)

compound
```

When adding property to compound and using a predefined PKML file, user need to also specify the 
corresponding path in the PKML model. 

### Compound Processes 

When automatically creating the generic models, the user can add processes to the compound but must 
use the addProcessProperty to do so and specify to which process type the parameter belongs to. 

```{r, eval = TRUE}
## add various processes to a compound by adding required process properties
compound$addProcessProperty(
  processType = "Liver Plasma Clearance",
  propertyName = "Plasma clearance",
  parName = "Plasma clearance",
  dimension = "Flow per weight",
  value = 0.27, 
  unit = "L/h/kg"
)

compound
```

## Formulations

Formulation can be defined using the various helper functions:
- `createDissolvedFormulation`
- `createWeibullFormulation`  
- `createLint80Formulation`
- `createParticleDissolutionFormulation`
- `createZeroOrderFormulation`
- `createFirstOrderFormulation`

```{r formulation, eval = TRUE}
weibull_form <- createWeibullFormulation(
  name = "Weibull",
  lagTime = 30,
  lagTimeUnit = "min",
  dissolutionTime50 = 60,
  dissolutionTime50Unit = "min"
)

weibull_form
```

## Administration protocols

The `SimpleProtocol` class is used to create the simple administration protocol that will be used 
in the studies. When a formulation is needed it can be associated with the protocol with 
`setFormulation`method

```{r, eval = TRUE}
protocol <- SimpleProtocol$new(
  route = "Oral",
  dosingInterval = "24",
  dose = 1,
  doseUnit = "mg/kg",
  startTime = 0,
  startTimeUnit = "h",
  endTime = 48,
  endTimeUnit = "h"
)

protocol$setFormulation(formulation = weibull_form)
protocol
```

Additionally advanced protocol can also be defined with the AdvancedProtocol class. To do so first 
an advanced protocol needs to be defined with the `AdvancedProtocol$new` function. Then schemas for repetitions need 
to be added with `addSchema` and single dose protocols can be added to the schema with 
`addProtocolToSchema` and will follow the schema repetition strategy.

```{r eval = TRUE}
protocol <- AdvancedProtocol$new()
protocol$addSchema(
  schemaName = "Loading Dose", 
  timeUnit = "h", 
  timeBetweenRepetitions = 0, 
  numberOfRepetitions = 1, 
  startTime = 0
)
protocol$addSchema(
  schemaName = "Maintenance Dose", 
  timeUnit = "h", 
  timeBetweenRepetitions = 12, 
  numberOfRepetitions = 5, 
  startTime = 12
)

ivLoading <- SimpleProtocol$new(route = "IV Infusion", dose = 10, doseUnit = "mg")
ivMaintenance <- SimpleProtocol$new(route = "IV Bolus", dose = 1, doseUnit = "mg")

protocol$addProtocolToSchema(schemaName = "Loading Dose", protocol = ivLoading)
protocol$addProtocolToSchema(schemaName = "Maintenance Dose", protocol = ivMaintenance)

protocol
```
