---
title: "Automatic Generic Models Creation with ESQhtpbpk"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Automatic Generic Models Creation with ESQhtpbpk}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 120
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The framework allows automatic detection of required processes and administrations, and creates the 
corresponding generic models.

First the package need to be loaded:

```{r setup}
library(ESQhtpbpk)
```

## Creating Studies

Next, the user needs to define simulation scenarios (i.e. studies) to be simulated. Each study 
encompasses every component needed for the simulation individual to be used, protocol and 
formulation for each compound administration and compound properties.

To simplify this task, various functions have been implemented in the framework easing the creation 
of all required parts of a study.

Additionally, the user can create its custom function to generate a list of all `Study` object to 
be simulated, tailored to their specific data structure, as this part is highly dependent on it.

### Example

The following example shows how to create a single study. This mechanism can be wrapped into a custom 
function to create a list of studies to be simulated using specific database as input for parameter value.

#### Compound Definition

First, the user needs to create a compound object with the `Compound$new()` function and assign a specific compound ID:

```{r comp_init, eval = TRUE}
compound <- Compound$new(ID = "Acetaminophen")
```

The compound is initialized with the main properties that are needed for the simulation. 
Those can easily be modified with the `setPropertyValue` method.

```{r comp_prop, eval = TRUE}
compound$setPropertyValue("Molecular weight", 151.16, unit = "g/mol")
compound$setPropertyValue("Solubility", 4.15, unit = "mg/ml")
compound$setPropertyValue("Lipophilicity", 0.91)
compound$setPropertyValue("pKa value 0", 9.46)
compound$setPropertyValue("Compound type 0", "Acidic")
compound$setPropertyValue("Fraction unbound", 80, unit = "%")
```

Additionally, various processes (as available in PK-Sim) can be added to the compound object using the `addProcessProperty` method. 
For example the following code adds a liver plasma clearance process by defining its plasma clearance:

```{r comp_add_process, eval = TRUE}
compound$addProcessProperty(
  processType = "Liver Plasma Clearance",
  propertyName = "Plasma clearance",
  parName = "Plasma clearance",
  dimension = "Flow per weight",
  value = 0.27, unit = "L/h/kg"
)
```

Additional properties for the same process can be added and their values adjusted if required in a similar fashion:

```{r comp_add_process_property, eval = TRUE}
compound$addProcessProperty(
  processType = "Liver Plasma Clearance",
  propertyName = "Lipophilicity",
  parName = "Lipophilicity (experiment)",
  dimension = "Log Units",
  value = 0.91
)
compound$addProcessProperty(
  processType = "Liver Plasma Clearance",
  propertyName = "Fraction unbound",
  parName = "Fraction unbound (experiment)",
  dimension = "Fraction",
  value = 80,
  unit = "%"
)
```

The partition coefficient and cellular permeability methods can be set easily with:

```{r choose_cp_pc, eval= TRUE}
compound$PartitionCoefficientMethod <- "PT"
compound$CellularPermeabilityMethod <- "PK-Sim"
```

The compound properties that have been defined can be easily visualized with:

```{r visualise_compound, eval = TRUE}
compound
```

#### Administration definition

Administration protocol and formulation can be easily defined as well.

The protocol can be defined with either `SimpleProtocol$new()` method for a simple protocol or
`AdvancedProtocol$new()` for a more complex protocol. Here we will use the `SimpleProtocol$new()` method:

```{r define_protocol, eval = TRUE}
# define protocol used in study and associate formulation used
protocol <- SimpleProtocol$new(
  route = "Oral",
  dosingInterval = "24",
  dose = 1,
  doseUnit = "mg/kg",
  startTime = 0,
  startTimeUnit = "h",
  endTime = 48,
  endTimeUnit = "h"
)        
```

For oral administration the formulation needs to be defined as well. Helper functions are available 
for each type of formulation and can be set as shown here for a Weibull formulation:

```{r define_formulation, eval = TRUE}
formulation <- createWeibullFormulation(
  name = "Weibull",
  lagTime = 30,
  lagTimeUnit = "min",
  dissolutionTime50 = 60,
  dissolutionTime50Unit = "min"
)
```

Then the formulation needs to be associated with the protocol defined previously using the `setFormulation` method.

```{r add_formulation_to_protocol, eval = TRUE}
protocol$setFormulation(formulation = formulation)
```

Similarly to compounds, administration protocol can be easily visualized:

```{r visualise_protocol, eval = TRUE}
protocol
```

Now that the protocol has been defined, it can be associated with the compound using the `setProtocol` method:

```{r add_protocol_to_compound, eval = TRUE}
compound$setProtocol(protocol)
```

#### Study definition

Finally, the study can be defined with the `Study$new()` method. The study needs to be given an unique ID:

```{r define_study, eval = TRUE}
study <- Study$new(ID = "Acetaminophen_PO_QD", compounds = list(compound), individual = "Human")
```

Similarly to the compound and protocol, the study can be easily visualized with:

```{r visualise_study, eval = TRUE}
study
```

## Running the studies

Once the user created the list of studies required, simulations can be run the using the `runPredictions()` function.
This function creates the required models, associated them with the corresponding studies and run the simulation.
All simulation generic models files are will be saved in the `outputFolder` within a time-stamped subfolder under GenericModels
and all simulations results can be saved under the same time-stamped subfolder under SimulationResults by setting the
`saveResults` parameter to `TRUE`. Additionally the user can save the PKML files for each study with the
`saveSimulation` parameter.

During this step the user should also defined the output paths to be simulated with the `outputSelections` parameter, as
well as the `simulationResolution` defined by `c(start time (min), end time (min), resolution in pts/min)` to cover the
longest simulation needed with the wanted time resolution.

Additionally the user can also defined the number of cores to be used for the simulations and the number studies to be
queued (which can be needed to reduce the memory usage).

```{r, eval = FALSE}
myresult <- runPredictions(
    studies = list(study), 
    outputFolder = "HT-PBPK",
    numberOfCores = 5,
    outputSelections = c("Organism|PeripheralVenousBlood|**|Plasma (Peripheral Venous Blood)"),
    simulationResolution = c(0, 10 * 24 * 60, 1),
    saveResults = TRUE,
    saveSimulation = FALSE,
    queueSize = 200
  )
```

## Current limitations

The current implementation does not allows:

-   user defined administration route
-   solubility table
-   formulations with dissolution tables
-   custom individual (only default individuals are supported)
-   gene expression use or definition
-   transport or enzymatic processes
-   food administration or other events
-   multiple solubility values at different pH
