---
title: "Automatic Generic Models Creation with ESQhtpbpk"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Automatic Generic Models Creation with ESQhtpbpk}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 120
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The framework allows automatic detection of needed process and administrations and creates the corresponding required
generic models.

First the package need to be loaded:

```{r setup}
library(ESQhtpbpk)
```

## Creating Studies

Then the user will need to define studies/scenarios that need to be simulated. These studies will recapitulate
everything that is needed for the simulation, individual to be used, protocol and formulation for each compound
administration and compound properties.

To this end, various functions have been implemented the ease the creation of all required parts of a study.

The user can create its custom function to generate a list of all `Study` object to be simulated, as this part is
dependent on the data structure or database used.

### Example

This example show how to create a single study. This mechanism could be wrapped into a function to create the list of
studies to be simulation and use database as input for parameter value assignment.

#### Compound definition

First the user needs to create a compound object with the `Compound$new()` function and set a specific compound ID:

```{r comp_init, eval = TRUE}
compound <- Compound$new(ID = "Acetaminophen")
```

The compound is initialized with the main properties that are needed for the simulation. Those can be easily adjusted
with the `setPropertyValue` method.

```{r comp_prop, eval = TRUE}
compound$setPropertyValue("Molecular weight", 151.16, unit = "g/mol")
compound$setPropertyValue("Solubility", 4.15, unit = "mg/ml")
compound$setPropertyValue("Lipophilicity", 0.91)
compound$setPropertyValue("pKa value 0", 9.46)
compound$setPropertyValue("Compound type 0", "Acidic")
compound$setPropertyValue("Fraction unbound", 80, unit = "%")
```

Additionally, various processes (as available in PK-Sim) can be added to the compound object. For example a liver plasma
clearance process can be added with the `addProcessProperty` method by defining its plasma clearance.

```{r comp_add_process, eval = TRUE}
compound$addProcessProperty(
  processType = "Liver Plasma Clearance",
  propertyName = "Plasma clearance",
  parName = "Plasma clearance",
  dimension = "Flow per weight",
  value = 0.27, unit = "L/h/kg"
)
```

Other properties of the same process can be added and their values adjusted if required in a similar fashion:

```{r comp_add_process_property, eval = TRUE}
compound$addProcessProperty(
  processType = "Liver Plasma Clearance",
  propertyName = "Lipophilicity",
  parName = "Lipophilicity (experiment)",
  dimension = "Log Units",
  value = 0.91
)
compound$addProcessProperty(
  processType = "Liver Plasma Clearance",
  propertyName = "Fraction unbound",
  parName = "Fraction unbound (experiment)",
  dimension = "Fraction",
  value = 80,
  unit = "%"
)
```

The partition coefficient method and cellular permeability method can be set easily with:

```{r choose_cp_pc, eval= TRUE}
compound$PartitionCoefficientMethod <- "PT"
compound$CellularPermeabilityMethod <- "PK-Sim"
```

The compound properties that have been defined can be easily visualised with:

```{r visualise_compound, eval = TRUE}
compound
```

#### Administration definition

Now that the compound has been defined the administration protocol and formulation can be easily defined as well.

The protocol can be easily defined with either `SimpleProtocol$new()` method for a simple protocol or
`AdvancedProtocol$new()` for a more complex protocol. Here we will use the `SimpleProtocol$new()` method:

```{r define_protocol, eval = TRUE}
# define protocol used in study and associate formulation used
protocol <- SimpleProtocol$new(
  route = "Oral",
  dosingInterval = "24",
  dose = 1,
  doseUnit = "mg/kg",
  startTime = 0,
  startTimeUnit = "h",
  endTime = 48,
  endTimeUnit = "h"
)        
```

When using an oral administration the formulation needs to be defined. Helper functions are available for each type of
formulation and can be set as shown here for a weibull formulation:

```{r define_formulation, eval = TRUE}
formulation <- createWeibullFormulation(
  name = "Weibull",
  lagTime = 30,
  lagTimeUnit = "min",
  dissolutionTime50 = 60,
  dissolutionTime50Unit = "min"
)
```

The formulation defined needs to be associated to the protocol already defined with the `setFormulation` method.

```{r add_formulation_to_protocol, eval = TRUE}
protocol$setFormulation(formulation = formulation)
```

Similarly to compound administration protocol can be easily visualised:

```{r visualise_protocol, eval = TRUE}
protocol
```

Now that the protocol is defined it can be associated to the compound with the `setProtocol` method:

```{r add_protocol_to_compound, eval = TRUE}
compound$setProtocol(protocol)
```

#### Study definition

Finally the study can be defined with the `Study$new()` method. The study needs to be given an unique ID,

```{r define_study, eval = TRUE}
study <- Study$new(ID = "Acetaminophen_PO_QD", compounds = list(compound), individual = "Human")
```

Similarly to the compound and protocol, the study can be easily visualised with:

```{r visualise_study, eval = TRUE}
study
```

## Running the studies

Once the user created the list of studies required, the user can run the simulation using the `runPredictions` function.
This function will create the required models, associated them with the corresponding studies and run the simulation.
All simulation generic models will be saved in the `outputFolder` within a time-stamped subfolder under GenericModels
and all simulations results can be saved under the same time-stamped subfolder under SimulationResults by setting the
`saveResults` parameter to `TRUE`. Additionally the user can save the PKML files for each study with the
`saveSimulation` parameter.

During this step the user should also defined the output paths to be simulated with the `outputSelections` parameter, as
well as the `simulationResolution` defined by `c(start time (min), end time (min), resolution in pts/min)` to cover the
longest simulation needed with the wanted time resolution.

Additionally the user can also defined the number of cores to be used for the simulations and the number studies to be
queued (which can be needed to reduce the memory usage).

```{r, eval = FALSE}
myresult <- runPredictions(
    studies = list(study), 
    outputFolder = "HT-PBPK",
    numberOfCores = 5,
    outputSelections = c("Organism|PeripheralVenousBlood|**|Plasma (Peripheral Venous Blood)"),
    simulationResolution = c(0, 10 * 24 * 60, 1),
    saveResults = TRUE,
    saveSimulation = FALSE,
    queueSize = 200
  )
```

## Current limitations

The current implementation does not allows:

-   user defined administration route
-   solubility table
-   formulations with dissolution tables
-   custom individual (only default individuals are supported)
-   gene expression use or definition
-   transport or enzymatic processes
-   food administration or other events
-   multiple solubility values at different pH
