---
title: "Using custom PKML models with ESQhtpbpk"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using custom PKML models with ESQhtpbpk}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The framework allows to use predefined pkml models for added/needed specificities.

First the package need to be loaded:

```{r setup}
library(ESQhtpbpk)
```

## Creating Studies

Then the user will need to define studies/scenarios that need to be simulated. These studies will recapitulate
everything that is needed for the simulation, individual to be used, protocol and formulation for each compound
administration and compound properties.

To this end, various functions have been implemented the ease the creation of all required parts of a study.

The user can create its custom function to generate a list of all `Study` object to be simulated, as this part is
dependent on the data structure or database used.

The main process is relatively similar to the one used for the automatic generation of generic models,
but when using predefined PKML models, the user will need to take care of specifying the correct path 
of every property to be updated based on his/her supplied PKML file.

### Example

This example show how to create a single study. This mechanism could be wrapped into a function to create the list of
studies to be simulation and use database as input for parameter value assignment.

The creation of studies is similar but paths to the given PKML files need to be provided.

#### Compound definition

First the user needs to create a compound object with the `Compound$new()` function and set a 
specific compound ID and the actual name of the compound to be updated from the PKML file (here 
`MyCompoundNameInPKML`)

```{r comp_init, eval = TRUE}
compound <- Compound$new(ID = "Acetaminophen", name = "MyCompoundNameInPKML")
```

The compound is initialized with the main properties that are needed for the simulation. Those can be easily adjusted
with the `setPropertyValue` method.

```{r comp_property, eval = TRUE} 
compound$setPropertyValue("Molecular weight", 151.16, unit = "g/mol")
compound$setPropertyValue("Solubility", 4.15, unit = "mg/ml")
compound$setPropertyValue("Lipophilicity", 0.91)
compound$setPropertyValue("pKa value 0", 9.46)
compound$setPropertyValue("Compound type 0", "Acidic")
compound$setPropertyValue("Fraction unbound", 80, unit = "%")
```

To add various processes related properties when using predefined PKML files, those can just be 
added with `addProperty` instead of `addProcessProperty` by giving the correct path in the 
predefined PKML model.

```{r comp_process_property, eval = TRUE} 
compound$addProperty(
  name = "THC - Plasma clearance", 
  path = "MyCompoundNameInPKML-Total Hepatic Clearance-Liver Plasma Clearance|Plasma clearance",
  parName = "THC - Plasma clearance",
  dimension = "Flow per weight",
  value = 0.27,
  unit = "L/h/kg"
)
compound$addProperty(
  name = "THC - Lipo", 
  path = "MyCompoundNameInPKML-Total Hepatic Clearance-Liver Plasma Clearance|Lipophilicity (experiment)",
  parName = "THC - Lipo",
  dimension = "Log Units",
  value = 0.91
)
compound$addProperty(
  name = "THC - Fu", 
  path = "MyCompoundNameInPKML-Total Hepatic Clearance-Liver Plasma Clearance|Fraction unbound (experiment)",
  parName = "THC - Fu",
  dimension = "Fraction",
  value = 80,
  unit = "%"
)
```

The defined compound can be visualized if needed :

```{r visualise_compound, eval = TRUE}
compound
```


The partition coefficient method and cellular permeability methods are set by default but not used 
in the case of a predefined PKML model, as it is already defined in the model and cannot be modified.

#### Administration definition

Similarly administration protocol can be defined but the corresponding path in the PKML model needs 
to be specified. For administration if multiple administrations are defined in the PKML model 
(for example 10 oral administrations are allowed in the PKML model), then the user should supplied 
all corresponding administration paths, only the needed ones will be used. 

```{r admin_init, eval = TRUE}
protocol <- SimpleProtocol$new(
  route = "Oral",
  dosingInterval = "24",
  dose = 1,
  doseUnit = "mg/kg",
  startTime = 0,
  startTimeUnit = "h",
  endTime = 48,
  endTimeUnit = "h", 
  path = paste0("Events|MyAdminNameInPKML|Application_", 1:10), 
)        
```

Formulations can also be defined the same way as for the automatic generation of generic models, 
but the user will need to supply the correct path to the formulation in the PKML model and associate
it to the protocol

```{r formulation, eval = T}
formulation <- createWeibullFormulation(
  name = "Weibull",
  lagTime = 30,
  lagTimeUnit = "min",
  dissolutionTime50 = 60,
  dissolutionTime50Unit = "min",
  path = "Events|MyAdminNameInPKML|MyFormNameInPKML",
)

protocol$setFormulation(formulation = formulation)
```

The defined protocol can be visualized if needed:

```{r visualize_protocol, eval = T}
protocol
```

The protocol can be associated to the compound
```{r associate_protocol, eval = TRUE}
compound$setProtocol(protocol)
```

The user should ensure that all the supplied paths are indeed consistent with the PKML model,
and if not correct them. 

```{r check_paths, eval = TRUE}
compound$getAllPropertyPaths()
```

#### Study definition

Finally the user can create the study object with the `Study$new()` function. The user will need to
supplied the path to the PKML file to be used for the simulation as well.

```{r example_study_creation_pkml, eval = FALSE} 
study <- Study$new(ID = "Acetaminophen_PO_QD", compounds = list(comp), 
                   genericModel = "path/to/your/pkml/MyGenericPKML.pkml")
```

Supplying individual is not needed in this case as it will use the one defined in the PKML file.

## Running the studies

Once the user created the list of studies required, the user can run the simulation using the
`runPredictions` function. This function will create the required models, associated them with 
the corresponding studies and run the simulation. All simulation generic models will be saved 
in the `outputFolder` within a time-stamped subfolder under GenericModels and all simulations 
results can be saved under the same time-stamped subfolder under SimulationResults by setting 
the `saveResults` parameter to `TRUE`. Additionally the user can save the PKML files for each 
study with the `saveSimulation` parameter.  

During this step the user should also defined the output paths to be simulated with the 
`outputSelections` parameter, as well as the `simulationResolution` defined by 
`c(start time (min), end time (min), resolution in pts/min)` to cover the longest 
simulation needed with the wanted time resolution.

Additionally the user can also defined the number of cores to be used for the simulations and 
the number studies to be queued (which can be needed to reduce the memory usage). 

### Example

```{r, eval = FALSE}
myresult <- runPredictions(
    studies = list(study), 
    outputFolder = "HT-PBPK",
    numberOfCores = 5,
    outputSelections = c("Organism|PeripheralVenousBlood|**|Plasma (Peripheral Venous Blood)"),
    simulationResolution = c(0, 10 * 24 * 60, 1),
    saveResults = TRUE,
    saveSimulation = FALSE,
    queueSize = 200
  )
```

