---
title: "Using custom PKML models with ESQhtpbpk"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using custom PKML models with ESQhtpbpk}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This framework enables to use predefined PKML models for for simulations with specific requirements.

First, the package need to be loaded:

```{r setup}
library(ESQhtpbpk)
```

## Creating Studies

Next, the user needs to define studies to be simulated. A study encompasses
everything that is required for the simulation, protocol and formulation for each compound
administration and compound properties.

To simplify this task, various functions have been implemented in the framework easing the creation 
of all required parts of a study.

Additionally, the user can create its custom function to generate a list of all `Study` object to 
be simulated, tailored to their specific data structure, as this part is highly dependent on it.

While the overall process is similar to the one used for the automatic generation of generic models,
when using predefined PKML models, the user must take care of specifying the correct path 
for each property to be updated based on his/her supplied PKML file.

### Example

This example shows how to create a single study. In practice, this mechanism can be wrapped into a 
function to create a list of all studies to be simulated, allowing integration with user databases 
or data structure for parameter assignment.

When working with predefined PKML files, the user must provide the correct paths for each property.

#### Compound Definition

First, the user creates a compound object using the `Compound$new()` function and assign a 
unique compound ID and gives the actual name of the compound to be updated from the predefined PKML 
file (here `MyCompoundNameInPKML`)

```{r comp_init, eval = TRUE}
compound <- Compound$new(ID = "Acetaminophen", name = "MyCompoundNameInPKML")
```

The compound is initialized with the main properties that are required for the simulation. 
Those can be easily adjusted with the `setPropertyValue` method.

```{r comp_property, eval = TRUE} 
compound$setPropertyValue("Molecular weight", 151.16, unit = "g/mol")
compound$setPropertyValue("Solubility", 4.15, unit = "mg/ml")
compound$setPropertyValue("Lipophilicity", 0.91)
compound$setPropertyValue("pKa value 0", 9.46)
compound$setPropertyValue("Compound type 0", "Acidic")
compound$setPropertyValue("Fraction unbound", 80, unit = "%")
```

To add additional processes related properties when using predefined PKML files, the user can simply
used the `addProperty` method (instead of `addProcessProperty`) while providing the correct path in the 
predefined PKML model.

```{r comp_process_property, eval = TRUE} 
compound$addProperty(
  name = "THC - Plasma clearance", 
  path = "MyCompoundNameInPKML-Total Hepatic Clearance-Liver Plasma Clearance|Plasma clearance",
  parName = "THC - Plasma clearance",
  dimension = "Flow per weight",
  value = 0.27,
  unit = "L/h/kg"
)
compound$addProperty(
  name = "THC - Lipo", 
  path = "MyCompoundNameInPKML-Total Hepatic Clearance-Liver Plasma Clearance|Lipophilicity (experiment)",
  parName = "THC - Lipo",
  dimension = "Log Units",
  value = 0.91
)
compound$addProperty(
  name = "THC - Fu", 
  path = "MyCompoundNameInPKML-Total Hepatic Clearance-Liver Plasma Clearance|Fraction unbound (experiment)",
  parName = "THC - Fu",
  dimension = "Fraction",
  value = 80,
  unit = "%"
)
```

The defined compound can be visualized for verification:

```{r visualise_compound, eval = TRUE}
compound
```


The partition coefficient and cellular permeability methods are set by default but not used 
in the case of a predefined PKML model, as it is already defined in the supplied model and cannot 
be modified.

#### Administration Definition

Administration protocol can be defined, but again, the corresponding path in the PKML model must 
be specified. 

If multiple administrations are defined in the PKML model (e.g. 10 oral 
administrations are allowed in the PKML model), then the user should supplied 
all relevant administration paths. Only the ones needed will be used in the simulation.

```{r admin_init, eval = TRUE}
protocol <- SimpleProtocol$new(
  route = "Oral",
  dosingInterval = "24",
  dose = 1,
  doseUnit = "mg/kg",
  startTime = 0,
  startTimeUnit = "h",
  endTime = 48,
  endTimeUnit = "h", 
  path = paste0("Events|MyAdminNameInPKML|Application_", 1:10), 
)        
```

Formulations can be defined the same way as for the automatic generation of generic models, 
but the user must supply the correct path to the formulation in the PKML model and also associate
it to the protocol. 

```{r formulation, eval = T}
formulation <- createWeibullFormulation(
  name = "Weibull",
  lagTime = 30,
  lagTimeUnit = "min",
  dissolutionTime50 = 60,
  dissolutionTime50Unit = "min",
  path = "Events|MyAdminNameInPKML|MyFormNameInPKML",
)

protocol$setFormulation(formulation = formulation)
```

The defined protocol can be visualized:

```{r visualize_protocol, eval = T}
protocol
```

The protocol is then be associated to the compound
```{r associate_protocol, eval = TRUE}
compound$setProtocol(protocol)
```

The user should make sure that all the supplied paths are indeed consistent with the PKML model,
and if not correct them before proceeding. 

```{r check_paths, eval = TRUE}
compound$getAllPropertyPaths()
```

#### Study definition

Finally, the user creates the study object with the `Study$new()` function. The user must supply 
the path to the PKML file to be used for the simulation.

```{r example_study_creation_pkml, eval = FALSE} 
study <- Study$new(ID = "Acetaminophen_PO_QD", compounds = list(comp), 
                   genericModel = "path/to/your/pkml/MyGenericPKML.pkml")
```

Note: An individual is not needed in this case, as it is already defined in the PKML file.

## Running the studies

Once the user created the list of studies to be simulation, the simulations can be run using the
`runPredictions` function. All simulations results can be saved under a `SimulationResults` 
time-stamped subfolder by setting the `saveResults` parameter to `TRUE`. Additionally the user 
can save the update PKML files for each study with the `saveSimulation` parameter.  

During this step the user should also defined the output paths to be simulated with the 
`outputSelections` parameter, as well as the `simulationResolution` defined by 
`c(start time (min), end time (min), resolution in pts/min)` to cover the longest 
simulation needed with the wanted time resolution.

Additionally the user can also defined the number of cores to be used for the simulations and 
the number studies to be queued (which can be needed to reduce the memory usage). 

```{r, eval = FALSE}
myresult <- runPredictions(
    studies = list(study), 
    outputFolder = "HT-PBPK",
    numberOfCores = 5,
    outputSelections = c("Organism|PeripheralVenousBlood|**|Plasma (Peripheral Venous Blood)"),
    simulationResolution = c(0, 10 * 24 * 60, 1),
    saveResults = TRUE,
    saveSimulation = FALSE,
    queueSize = 200
  )
```

