---
title: "Implementation details"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Implementation details}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library("ESQhtpbpk")
```


## Generic PBPK model structure

A generic PBPK model will be automatically created for each species, partition coefficient (PC) method, 
cellular permeability (CP) method, as well as each required combination of processes for the number of 
compounds used. For each generic model structure the required number of administrations of each type 
covering all studies using this generic model will be identified and added.

Note: CP calculation method "Charge dependent Schmitt normalized to
PK-Sim" is not supported as it is not properly implemented in the
current PK-Sim versions (11.3 and 12.0).

## Compound

The `Compound` class is used to create the compound object that will be used in the studies.
Several compound properties are already predefined and can be directly used. 

The predefined compound properties (with their default values) are:

```{r, eval = TRUE}
compound <- Compound$new(ID = "Acetaminophen")
compound
```

### Compound Properties

Predefined compound properties can be easily changed with the `setPropertyValue` method

```{r, eval = TRUE}
compound$setPropertyValue("Molecular weight", 151.16, unit = "g/mol")
```

Additional compound properties can be added as needed using `addProperty` methods (with corresponding paths when using 
predefined PKML files). The property `name` serves as an alias for easy access in the compound object. The `parName` 
must match the parameter name as defined in PK-Sim. 

```{r, eval = TRUE}
compound$addProperty(
  name = "Pint", 
  parName = "Intestinal permeability (transcellular)",
  dimension = "Velocity",
  value = 2e-5,
  unit = "cm/min"
)

compound
```

When adding property to compound and using a predefined PKML file, the user also needs to specify the 
corresponding path in the PKML model. 

### Compound Processes 

When creating generic models automatically, the user can add processes to the compound using the 
`addProcessProperty` method. This requires to specify to which process type the parameter belongs to. 

```{r, eval = TRUE}
## add various processes to a compound by adding required process properties
compound$addProcessProperty(
  processType = "Liver Plasma Clearance",
  propertyName = "Plasma clearance",
  parName = "Plasma clearance",
  dimension = "Flow per weight",
  value = 0.27, 
  unit = "L/h/kg"
)

compound
```

## Formulations

Formulations can be defined using the various helper functions provided:
- `createDissolvedFormulation`
- `createWeibullFormulation`  
- `createLint80Formulation`
- `createParticleDissolutionFormulation`
- `createZeroOrderFormulation`
- `createFirstOrderFormulation`

```{r formulation, eval = TRUE}
weibull_form <- createWeibullFormulation(
  name = "Weibull",
  lagTime = 30,
  lagTimeUnit = "min",
  dissolutionTime50 = 60,
  dissolutionTime50Unit = "min"
)

weibull_form
```

## Administration protocols

The `SimpleProtocol` class is used to define simple administration protocols for the studies. 
If a formulation is needed it can be associated with the protocol using the `setFormulation` method.

```{r, eval = TRUE}
protocol <- SimpleProtocol$new(
  route = "Oral",
  dosingInterval = "24",
  dose = 1,
  doseUnit = "mg/kg",
  startTime = 0,
  startTimeUnit = "h",
  endTime = 48,
  endTimeUnit = "h"
)

protocol$setFormulation(formulation = weibull_form)
protocol
```

Additionally, for more complex dosing strategies, advanced protocols can be created with the 
`AdvancedProtocol` class. To do so, first an advanced protocol needs to be initialized with the 
`AdvancedProtocol$new` function. Then repetitions schemas need to be defined with `addSchema` method
and associated with single dose protocol using `addProtocolToSchema` method. 
These will follow the schema repetition strategy defined during the `addSchema` step.

```{r eval = TRUE}
protocol <- AdvancedProtocol$new()
protocol$addSchema(
  schemaName = "Loading Dose", 
  timeUnit = "h", 
  timeBetweenRepetitions = 0, 
  numberOfRepetitions = 1, 
  startTime = 0
)
protocol$addSchema(
  schemaName = "Maintenance Dose", 
  timeUnit = "h", 
  timeBetweenRepetitions = 12, 
  numberOfRepetitions = 5, 
  startTime = 12
)

ivLoading <- SimpleProtocol$new(route = "IV Infusion", dose = 10, doseUnit = "mg")
ivMaintenance <- SimpleProtocol$new(route = "IV Bolus", dose = 1, doseUnit = "mg")

protocol$addProtocolToSchema(schemaName = "Loading Dose", protocol = ivLoading)
protocol$addProtocolToSchema(schemaName = "Maintenance Dose", protocol = ivMaintenance)

protocol
```
