---
title: "HT-PBPK Prediction Pipeline"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{HT-PBPK Prediction Pipeline}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The high-throughput (HT) physiologically-based (PB) pharmacokinetics
(PK) modeling framework is distributed as an R package that allow to predict 
the pharmacokinetics of various molecules in the various species. It uses the capability 
of OSP Suite PK-Sim 

# Main Features

- Allow to run efficiently and in parallel many compound prediction.
- Can automatically create the various generic models required (limited to PK-Sim components).
- Can be used with custom pkmls for added/needed specificities.

# Current limitations (for the automatic generic models creation)

The current implementation does not allows:
- user defined administration route
- solubility table
- formulations with dissolution tables
- custom individual (only default individuals are supported)
- gene expression use or definition
- transport or enzymatic processes
- food administration or other events
- multiple solubility values at different pH


# Running the framework

For the package need to be loaded:

```{r setup}
library(ESQhtpbpk)
```

## Creating Studies

There is to main options to run the framework either by using the automatic generation of models 
or by using custom pkmls.

These two options will be described separately as their requirements are 
slightly different.

### Using the automatic generation of generic models

The user will need to define studies/scenarios that need to be simulated. This studies will 
recapitulate everything that is needed for the simulation, individual to be used, protocol 
and formulation for each compound administration, compound properties.

`Compounds` object and various functions have been implemented the ease the creation of all 
required parts of a study for more details see `Implementation details` section. 

The user can create its custom function to generate a list of all `Study` object to be simulated, 
as this part is dependent on the data structure or database used.

#### Example

This example show how to create a single study. This mechanism could be wrapped into a function
to create the list of studies to be simulation and use database as input for parameter value 
assignment. 

```{r example_study_creation, eval = TRUE} 
# create compounds object
comp <- Compound$new(ID = "Acetaminophen")

## set various compound properties
comp$setPropertyValue("Molecular weight", 151.16, unit = "g/mol")
comp$setPropertyValue("Solubility", 4.15, unit = "mg/ml")
comp$setPropertyValue("Lipophilicity", 0.91)
comp$setPropertyValue("pKa value 0", 9.46)
comp$setPropertyValue("Compound type 0", "Acidic")
comp$setPropertyValue("Fraction unbound", 80, unit = "%")

## add various processes to a compound by adding required process properties
comp$addProcessProperty(
  processType = "Liver Plasma Clearance",
  propertyName = "Plasma clearance",
  parName = "Plasma clearance",
  dimension = "Flow per weight",
  value = 0.27, unit = "L/h/kg"
)
comp$addProcessProperty(
  processType = "Liver Plasma Clearance",
  propertyName = "Lipophilicity",
  parName = "Lipophilicity (experiment)",
  dimension = "Log Units",
  value = 0.91
)
comp$addProcessProperty(
  processType = "Liver Plasma Clearance",
  propertyName = "Fraction unbound",
  parName = "Fraction unbound (experiment)",
  dimension = "Fraction",
  value = 80,
  unit = "%"
)

# choose CP/PC method to be used for compound
comp$PartitionCoefficientMethod <- "PT"
comp$CellularPermeabilityMethod <- "PK-Sim"

## defined compound can be visualised if needed
comp

# define formulation needed in the protocol
formulation <- createWeibullFormulation(
  name = "Weibull",
  lagTime = 30,
  lagTimeUnit = "min",
  dissolutionTime50 = 60,
  dissolutionTime50Unit = "min"
)

# define protocol used in study and associate formulation used
protocol <- SimpleProtocol$new(
  route = "Oral",
  dosingInterval = "24",
  dose = 1,
  doseUnit = "mg/kg",
  startTime = 0,
  startTimeUnit = "h",
  endTime = 48,
  endTimeUnit = "h"
)        

protocol$setFormulation(formulation = formulation)

## defined protocol can be visualised if needed
protocol

# associate protocol to compound
comp$setProtocol(protocol)

# define study to be simulated
study <- Study$new(ID = "Acetaminophen_PO_QD", compounds = list(comp), individual = "Human")
```

### Using the predefined pkmls

The user will need to define studies/scenarios that need to be simulated. This studies will 
recapitulate everything that is needed to be changed for the simulation, protocol and formulation 
need for each compound with corresponding path as defined in the simulation, compounds properties 
with corresponding paths in the simulation and the pkml model to be used.

The user can create its custom function to generate a list of all `Study` object to be simulated, 
as this part is dependent on the data structure or database used.

#### Example
The creation of studies is similar but paths to the given pkmls need to be provided.
```{r example_study_creation_pkml, eval = TRUE} 
# create compounds object name must correspond to the compound name in the wanted pkml
comp <- Compound$new(ID = "Acetaminophen", name = "MyCompoundNameInPKML")

## set various compound properties
comp$setPropertyValue("Molecular weight", 151.16, unit = "g/mol")
comp$setPropertyValue("Solubility", 4.15, unit = "mg/ml")
comp$setPropertyValue("Lipophilicity", 0.91)
comp$setPropertyValue("pKa value 0", 9.46)
comp$setPropertyValue("Compound type 0", "Acidic")
comp$setPropertyValue("Fraction unbound", 80, unit = "%")

## add various processes, for predefined pkml processes can just be added with addProperty instead 
## of addProcessProperty by giving the correct path
comp$addProperty(
  name = "THC - Plasma clearance", 
  path = "MyCompoundNameInPKML-Total Hepatic Clearance-Liver Plasma Clearance|Plasma clearance",
  parName = "THC - Plasma clearance",
  dimension = "Flow per weight",
  value = 0.27,
  unit = "L/h/kg"
)
comp$addProperty(
  name = "THC - Lipo", 
  path = "MyCompoundNameInPKML-Total Hepatic Clearance-Liver Plasma Clearance|Lipophilicity (experiment)",
  parName = "THC - Lipo",
  dimension = "Log Units",
  value = 0.91
)
comp$addProperty(
  name = "THC - Fu", 
  path = "MyCompoundNameInPKML-Total Hepatic Clearance-Liver Plasma Clearance|Fraction unbound (experiment)",
  parName = "THC - Fu",
  dimension = "Fraction",
  value = 80,
  unit = "%"
)

## defined compound can be visualised if needed (PC/CP methods are set by default but not used in 
## the case of a supplied pkml)
comp

# define formulation needed in the protocol with correct path from the pkml
formulation <- createWeibullFormulation(
  name = "Weibull",
  lagTime = 30,
  lagTimeUnit = "min",
  dissolutionTime50 = 60,
  dissolutionTime50Unit = "min",
  path = "Events|MyAdminNameInPKML|MyFormNameInPKML",
)

# define protocol used in study with all usable admin path of the same type if multiple 
# administrations are defined in the pkml (for example 10 oral administrations are allowed in the 
# pkml, you can supplied all corresponding administration paths, only the needed ones will be used)
protocol <- SimpleProtocol$new(
  route = "Oral",
  dosingInterval = "24",
  dose = 1,
  doseUnit = "mg/kg",
  startTime = 0,
  startTimeUnit = "h",
  endTime = 48,
  endTimeUnit = "h", 
  path = paste0("Events|MyAdminNameInPKML|Application_", 1:10), 
)        

protocol$setFormulation(formulation = formulation)

## defined protocol can be visualised if needed
protocol

# associate protocol to compound
comp$setProtocol(protocol)

# all supplied path to be replaced can be checked for consistency with the pkml
comp$getAllPropertyPaths()
```

```{r example_study_creation_pkml2, eval = FALSE} 
# define study to be simulated, when using a predefined pkml, the individual is not needed
# as it will be the one defined in the pkml.
study <- Study$new(ID = "Acetaminophen_PO_QD", compounds = list(comp), 
                   genericModel = "path/to/your/pkml/MyGenericPKML.pkml")
```

### Running the studies

Once the user created the list of studies required, the user can run the simulation using the
`runPredictions` function. This function will create the required models, associated them with 
the corresponding studies and run the simulation. All simulation generic models will be saved 
in the `outputFolder` within a time-stamped subfolder under GenericModels and all simulations 
results can be saved under the same time-stamped subfolder under SimulationResults by setting 
the `saveResults` parameter to `TRUE`. Additionally the user can save the pkmls for each study 
with the `saveSimulation` parameter.  

During this step the user should also defined the output paths to be simulated with the 
`outputSelections` parameter, as well as the `simulationResolution` defined by 
`c(start time (min), end time (min), resolution in pts/min)` to cover the longest 
simulation needed with the wanted time resolution.

Additionally the user can also defined the number of cores to be used for the simulations and 
the number studies to be queued (which can be needed to reduce the memory usage). 

#### Example

```{r, eval = FALSE}
myresult <- runPredictions(
    studies = list(study), 
    outputFolder = "HT-PBPK",
    numberOfCores = 5,
    outputSelections = c("Organism|PeripheralVenousBlood|**|Plasma (Peripheral Venous Blood)"),
    simulationResolution = c(0, 10 * 24 * 60, 1),
    saveResults = TRUE,
    saveSimulation = FALSE,
    queueSize = 200
  )
```

### Metrics - Not Implemented yet

If observed datasets are available for a specific study, comparison metrics will be calculated 
automatically by the runPrediction function and saved.

# Implementation details

## Generic PBPK model structure

A generic PBPK model will be automatically created for each species, partition coefficient method (PC), 
cellular permeability method (CP) as well as each required combination of processes for a certain number 
of compounds used. For each generic model structure the required administrations from all studies 
using this generic model will be checked and added.

Note: CP calculation method "Charge dependent Schmitt normalized to
PK-Sim" is not supported as it is not properly implemented in the
current PK-Sim versions (11.3 and 12.0).

## Compound

The `Compound` class is used to create the compound object that will be used in the studies.
Some compound properties are already predefined and can be directly used. 

Pre-specified compound properties (with their default values) are:

```{r, eval = TRUE}
comp <- Compound$new(ID = "Acetaminophen")
comp
```
### Compound Properties

Additional compound properties can be added as needed with addProperty (with corresponding paths for 
predefined pkmls). Existing ones can be easily changed with the setPropertyValue function.
The property `name` is an alias to easily access a property in the compound object. The `parName` 
needs to correspond to the parameter name as defined in the PK-Sim. 

When adding property to compound and using a predefined pkml, user need to specify the 
corresponding path in the pkml. 

```{r, eval = TRUE}
comp$setPropertyValue("Molecular weight", 151.16, unit = "g/mol")

comp$addProperty(
  name = "Pint", 
  parName = "Intestinal permeability (transcellular)",
  dimension = "Velocity",
  value = 2e-5,
  unit = "cm/min"
)

comp
```
### Compound Processes 

When automatically creating the generic models, the user can add processes to the compound but must 
use the addProcessProperty to do so and specifiy to which process type the parameter belongs to. 

```{r, eval = TRUE}
## add various processes to a compound by adding required process properties
comp$addProcessProperty(
  processType = "Liver Plasma Clearance",
  propertyName = "Plasma clearance",
  parName = "Plasma clearance",
  dimension = "Flow per weight",
  value = 0.27, 
  unit = "L/h/kg"
)

comp
```

## Formulations

Formulation can be defined using the various helper functions:
- `createDissolvedFormulation`
- `createWeibullFormulation`  
- `createLint80Formulation`
- `createParticleDissolutionFormulation`
- `createZeroOrderFormulation`
- `createFirstOrderFormulation`

```{r formulation, eval = TRUE}
form_weib <- createWeibullFormulation(
  name = "Weibull",
  lagTime = 30,
  lagTimeUnit = "min",
  dissolutionTime50 = 60,
  dissolutionTime50Unit = "min"
)

form_weib
```

## Administration protocols

The `SimpleProtocol` class is used to create the simple administration protocol that will be used 
in the studies. When a formulation is needed it can be associated with the protocol with setFormulation

```{r, eval = TRUE}
prot <- SimpleProtocol$new(
  route = "Oral",
  dosingInterval = "24",
  dose = 1,
  doseUnit = "mg/kg",
  startTime = 0,
  startTimeUnit = "h",
  endTime = 48,
  endTimeUnit = "h"
)

prot$setFormulation(formulation = form_weib)
prot
```

Additionally advanced protocol can also be defined with the AdvancedProtocol class. To do so first 
an advanced protocol needs to be defined with the `AdvancedProtocol$new` function. Then schemas for repetitions need 
to be added with `addSchema` and single dose protocols can be added to the schema with 
`addProtocolToSchema` and will follow the schema repetition strategy.

```{r eval = TRUE}
prot <- AdvancedProtocol$new()
prot$addSchema(
  schemaName = "Loading Dose", 
  timeUnit = "h", 
  timeBetweenRepetitions = 0, 
  numberOfRepetitions = 1, 
  startTime = 0
)
prot$addSchema(
  schemaName = "Maintenance Dose", 
  timeUnit = "h", 
  timeBetweenRepetitions = 12, 
  numberOfRepetitions = 5, 
  startTime = 12
)

ivLoading <- SimpleProtocol$new(route = "IV Infusion", dose = 10, doseUnit = "mg")
ivMaintenance <- SimpleProtocol$new(route = "IV Bolus", dose = 1, doseUnit = "mg")

prot$addProtocolToSchema(schemaName = "Loading Dose", protocol = ivLoading)
prot$addProtocolToSchema(schemaName = "Maintenance Dose", protocol = ivMaintenance)

prot
```
