---
title: "Example with chembl dataset"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r, include = FALSE}
# create custom DT for inputs 
myTinyDT <- function(df) {
  DT::datatable(
    df, 
    options = list(
      scrollX = T, 
      lengthChange = FALSE,
      pageLength = 5
    ),
    class = "compact",
    rownames = FALSE
  ) %>% 
  DT::formatStyle(
    columns = names(df),
    fontSize = "12px", 
    "white-space" = "nowrap"
  ) %>%
  htmltools::tagList(
    htmltools::tags$style(
      htmltools::HTML(
        "table.dataTable {font-size: 12px; white-space:nowrap}",
        ".dataTables_wrapper {font-size: 12px}"
      )
    )
  )
}
```

## Load needed libraries

For this example we will load a few libraries

```{r setup, echo = TRUE, warning = FALSE, message = FALSE}
library(ESQhtpbpk) # package for HTPBPK
library(readxl) # package to read xlsx files that contains input data
library(dplyr) # package for easy manipulation of data
library(tidyr) # package for easy manipulation of data
library(stringr) # package for string manipulation
# packages for plotting
library(ggplot2)
library(ggrepel)
library(scales)
library(plotly)
# packages for parallel computing
library(doParallel)
```


## Load input files

A dataset of Cmax and AUC extracted and curated from CHEMBL is available in the package as an example dataset.

```{r loading_datasets}
# Reading curated chembl dataset with AUC and Cmax values
path <- system.file(
  "extdata", 
  "Chembl_preprocessed_curated.xlsx", 
  package = "ESQhtpbpk"
)
```

The first sheet of the excel list the studies and the observed data available for each study.

```{r loading_studies}
# Reading curated chembl dataset with AUC and Cmax values
InputStudies <- read_excel(path, sheet = 1)
```


```{r echo = FALSE}
myTinyDT(InputStudies)
```

The second sheet of the excel list the protocols information to be simulated. A protocol can spread 
multiple lines for the same study for example in case of multiple administration with loading dose. 

```{r loading_protocol}
InputProtocol <- read_excel(path, sheet = 2)
```  

```{r echo = FALSE}
myTinyDT(InputProtocol)
```

The third sheet of the excel list all the compound properties. Based on 
[this paper](https://link.springer.com/article/10.1007/s00204-024-03764-9) and more specifically 
[Table 1](https://link.springer.com/article/10.1007/s00204-024-03764-9/tables/1), we extracted 
compound properties from [ADMETlab](https://admetlab3.scbdd.com/) QSAR as input to the HT-PBPK model. 

```{r loading_compounds}
InputCompounds <- read_excel(path, sheet = 3) # QSARs obtained from ADMETlab 3.0
```

```{r echo = FALSE}
myTinyDT(InputCompounds)
```

### Summary of compound properties included in the dataset

This dataset was extracted from Chembl and further curated to contain only studies with a single compound administered in healthy adults for which AUC or Cmax was available and protocol was identified from the original paper.

With the careful curation, the dataset ends up only containing `r length(unique(InputStudies$molecule_chembl_id))` compounds.  

Here is a summary of the properties for compounds included:

```{r compounds_summary, message = FALSE}
InputCompounds %>% 
  pivot_longer(
    cols = c(MW, logP, logD, logS, Fu, PPB, MDCK, caco2, PAMPA, 
             `cl-plasma`, pka_acidic, pka_basic),
    names_to = "Property", 
    values_to = "Value"
  ) %>% 
  ggplot() +
    geom_histogram(aes(x = Value)) + 
    facet_wrap(~ Property, scales = "free")
``` 


## Testing multiple scenarios

The platform allows to test easily multiple scenarios by adjusting inputs as needed based on the chosen scenarios. 
An example of excel file with different scenarios is available in the package. 
For example, the solubility can be set to an arbitrary high value or to the value from QSAR, the 
lipophilicity can be set to logP or logD, etc.

```{r load_scenarios}
scenarios <- read.csv(
  file = system.file(
    "extdata", 
    "Chembl_Test_Pipeline", 
    "scenarios.csv", 
    package = "ESQhtpbpk"
  ), 
  stringsAsFactors = FALSE
)
```

```{r echo = FALSE}
myTinyDT(scenarios)
```

The various scenarios can be easily implemented when setting up the compound by using 
if or switch condition when setting the inputs (see below).  

## Running the pipeline 

First we can set the outputFolder where we will save the results.

```{r setFolder, eval = F}
outputFolder <- "Chembl_Test_Pipeline"
dir.create(outputFolder, showWarnings = FALSE)
```

To create studies, it will depends on how our inputs are formatted, thus it can be very useful to create custom functions that are tailored to our inputs.

### Compound creation for various scenarios 

First let's create a custom function that will create a compound object based on our inputs and scenario tested, then it will be easy to wrap it to create all the study we need for all the scenarios we want.

```{r create_study_function, eval = T}
createCompound <- function(chembl_id, scenario) {
  compoundIdx <- which(InputCompounds$molecule_chembl_id == chembl_id)

  # Set new compound with compound name corresponding to chembl_id
  comp <- Compound$new(ID = chembl_id)
  
  # if halogens are considered set halogenes numbers
  if (scenario$Halogen == 1) {
    # Set halogen numbers (up to 10 allowed in PK-Sim)
    comp$setPropertyValue(
      "Chlorine count", 
      value = min(10, str_count(InputCompounds$smiles[compoundIdx], "Cl"))
    )
    comp$setPropertyValue(
      "Bromine count", 
      value = min(10, str_count(InputCompounds$smiles[compoundIdx], "Br"))
    )
    comp$setPropertyValue(
      "Fluorine count", 
      value = min(10, str_count(InputCompounds$smiles[compoundIdx], "F"))
    )
    comp$setPropertyValue(
      "Iodine count", 
      value = min(10, str_count(InputCompounds$smiles[compoundIdx], "I"))
    )
  }
    
  # set molecular weight
  comp$setPropertyValue(
    "Molecular weight", 
    value = InputCompounds$MW[compoundIdx], 
    unit = "g/mol"
  )
    
  # Set solubility depending on scenario
  comp$setPropertyValue(
    "Solubility", 
    value = switch(
      as.character(scenario$Solubility), 
      # set solubility from QSAR
      # ADMETlab logS QSAR is given as log10 of mol/l, convert to g/l for PK-Sim
      logS = 10 ^ InputCompounds$logS[compoundIdx] * InputCompounds$MW[compoundIdx], 
      # set arbitrary high solubility
      High = 1000
    ),
    unit = "g/l"
  )
    
  # Set lipophilicity based on scenario
  comp$setPropertyValue(
    "Lipophilicity", 
    value = switch(
      as.character(scenario$Lipophilicity), 
      # set  lipophilicity to logP
      logP = InputCompounds$logP[compoundIdx],
      # set  lipophilicity to logD
      logD = InputCompounds$logD[compoundIdx],
      # use logMA as lipophilicity input
      logMA1 = 1.294 + 0.304 * InputCompounds$logP[compoundIdx],
      # use logMA as lipophilicity input
      logMA_Pearce = 0.0166 * 37 + 0.882 * InputCompounds$logP[compoundIdx]
    ), 
    unit = "Log Units"
  )
    
  # Set Fu based on scenario
  comp$setPropertyValue(
    "Fraction unbound", 
    value = switch(
      as.character(scenario$Fu), 
      # use Fu QSAR
      Fu = InputCompounds$Fu[compoundIdx],
      # use PPB QSAR
      PPB = 100 - InputCompounds$PPB[compoundIdx]
    ), 
    unit = "%"
  )

  # Set intestinal permeability based on scenario 
  # intestinal permeability is not predefined it needs to be added with addProperty
  if (scenario$Pint != "PK-Sim") {
    comp$addProperty(
      name = "PInt",
      parName = "Specific intestinal permeability (transcellular)",
      dimension = "Velocity",
      value = switch(
        as.character(scenario$Pint), 
        # set Pint to MDCK QSAR value
        MDCK = 10^InputCompounds$MDCK[compoundIdx], # MDCK is in log cm/s,
        # set Pint to Caco2 QSAR value
        Caco2 = 10^InputCompounds$caco2[compoundIdx], # Caco2 is in log cm/s,
        # set Pint to PAMPA QSAR value
        PAMPA = InputCompounds$PAMPA[compoundIdx] * 1e-6 # PAMPA is in 10E-6 cm/s,
      ),
      unit = "cm/s"
    )
  }
  
  # Set organ permeability to arbitrarily high value if scenario requires it
  if (scenario$Perm ==  "High") {
    comp$addProperty(
      name = "Perm",
      parName = "Permeability",
      dimension = "Velocity",
      value = 10,
      unit = "cm/s"
    )
  }
   
  # Add total clearance (as hepatic clearance) also add matching fu and lipophilicity
  comp$addProcessProperty(
    processType = "Liver Plasma Clearance",
    propertyName = "Plasma clearance",
    parName = "Plasma clearance",
    dimension = "Flow per weight",
    value = InputCompounds$`cl-plasma`[compoundIdx], 
    unit = "ml/min/kg"
  )
  comp$addProcessProperty(
    processType = "Liver Plasma Clearance",
    propertyName = "Lipophilicity",
    parName = "Lipophilicity (experiment)",
    dimension = "Log Units",
    value = suppressMessages(comp$getProperty("Lipophilicity")$value),
    unit = suppressMessages(comp$getProperty("Lipophilicity")$unit)
  )
  comp$addProcessProperty(
    processType = "Liver Plasma Clearance",
    propertyName = "Fraction unbound",
    parName = "Fraction unbound (experiment)",
    dimension = "Fraction",
    value = suppressMessages(comp$getProperty("Fraction unbound")$value),
    unit = suppressMessages(comp$getProperty("Fraction unbound")$unit)
  )
    
  # Consider pka depending on scenario (restrict to 0-14 range)
  if (scenario$PkA == 1) {
    comp$setPropertyValue(
      name = "pKa value 0", 
      value = min(max(c(0,InputCompounds$pka_acidic[compoundIdx])), 14)
    )
    comp$setPropertyValue("Compound type 0", "Acidic")
      
    comp$setPropertyValue(
      name = "pKa value 1", 
      value = min(max(c(0,InputCompounds$pka_basic[compoundIdx])), 14)
    )
    comp$setPropertyValue("Compound type 1", "Basic")
  }
  
  # Include default GFR fraction of 1 based on scenario
  if (scenario$GFR == 1) {
    comp$addProcessProperty(
      processType = "GFR",
      propertyName = "GFR",
      parName = "GFR fraction",
      dimension = "Dimensionless",
      value = 1
    )
  }

  # Set liver enzyme concentration to 1 umol/l in 
  # intracellular space if enzyme correction tested
  if (scenario$Enzyme_correction == 1) {
    comp$addProcessProperty(
      processType = "Liver Plasma Clearance",
      propertyName = "Enzyme correction",
      parName = "Enzyme concentration",
      dimension = "Concentration (molar)",
      value = 1/0.667, # 1/0.667 is the correction factor for the enzyme activity
      unit = "umol/l"
    )
  }
  
  return(comp)
}
```


### Protocol creation 

We can also do a custom function to create our administration protocols based on the way our protocol information are set up.

```{r create_protocol_function, eval = TRUE}
createProtocol <- function(protocol_id, scenario) {
  # extract lines corresponding to the wanted protocol
  protocolTable <- InputProtocol[InputProtocol$ProtocolID == protocol_id,]
      
  # create advanced protocol (1 protocol can be defined by multiple lines
  # in the excel with different admin, for example with loading dose)
  prot <- AdvancedProtocol$new()
  
  # skip if some dose is not available (look depending on scenario selected)
  doseVal <- switch(
    as.character(scenario$Dose), 
    Flat = protocolTable$`Dose`,
    BW = protocolTable$`Dose Norm`
  )
  doseUnit <- switch(
    as.character(scenario$Dose), 
    Flat = protocolTable$`Dose Unit`,
    BW = protocolTable$`Dose Norm Unit`
  )

  if (any(is.na(doseVal) | is.na(doseUnit))) {
    return(NULL)
  }
  
  # Add schema for each line of the protocol info    
  for (i in seq_len(nrow(protocolTable))) {
    # default is 1 dose, with 0 time between repetitions if not specified
    prot$addSchema(
      schemaName = paste0("Schema_", i),
      startTime = protocolTable$Time[i],
      timeBetweenRepetitions = protocolTable$Time_between_rep[i] %>% ifelse(is.na(.), 0, .),
      numberOfRepetitions = protocolTable$Nrep[i] %>% ifelse(is.na(.), 1, .),
      timeUnit = protocolTable$TimeUnit[i]
    )
    
    # select dose type based on scenario (dose as given, or always converted to per kg)
    admin <- SimpleProtocol$new(
      # match route from excel to allowed route in package
      route = c("ORAL" = "Oral", "IV INFUSION" = "IV Infusion")[protocolTable$Route[i]],
      dose = doseVal[i], 
      doseUnit = doseUnit[i], 
      infusionTime = (protocolTable$Infusion_duration[i] %>% if (is.na(.)) NULL else .), 
      infusionTimeUnit = (protocolTable$Infusion_duration_unit[i] %>% if (is.na(.)) NULL else .)
    )
    
    # for oral admin assume dissolved formulation      
    if (protocolTable$Route[i] == "Oral") {
      formulation <- createDissolvedFormulation(name = "Dissolved")
      admin$setFormulation(formulation = formulation)
    }
    
    # add schema to protocol    
    prot$addProtocolToSchema(admin, schemaName = paste0("Schema_", i))
  }
  
  return(prot)
}
``` 

### Studies creation

Then we can easily use these function to create the list of studies for each scenario we want to test

```{r create_studies, eval = TRUE}
createStudies <- function(scenario) {
  
  # initialise to an empty list of studies
  Studies <- list()

  # we loop over all compounds (for faster execution, 
  # as that way the compound will need to be created only once)
  # we could also loop over the studies but then for each study 
  # the compound would need to be created again
  
  for (compoundIdx in seq_len(nrow(InputCompounds))) {
    # create compound with inputs matching wanted scenario
    compChemblId <- InputCompounds$molecule_chembl_id[compoundIdx]
    
    comp <- createCompound(
      chembl_id = compChemblId, 
      scenario = scenario
    )
    
    # create studies using the compound (as in our case studies only use a single compound)
    # our input data as 1 ine per observation hence the use of unique function
    studyIdsForComp <- unique(
      InputStudies$StudyID[which(InputStudies$molecule_chembl_id == compChemblId)]
    )
    for (studyId in studyIdsForComp) {
      print(studyId)
      studyIdx <- which(InputStudies$StudyID == studyId)
      
      # create protocol for the study
      prot <- createProtocol(
        protocol_id = unique(InputStudies$ProtocolID[studyIdx]), 
        scenario = scenario
      )
      
      if (is.null(prot)) {
        # skip if no protocol available (i.e. missing dose information 
        # for example for BW norm dose if BW was not available)
        next()
      }
      
      # set protocol to the compound
      comp$setProtocol(prot)
      
      # set wanted PC and CP method
      comp$PartitionCoefficientMethod <- "PK-Sim"
      comp$CellularPermeabilityMethod <- "PK-Sim"
          
      # create study with the compound and protocol for human individual
      study <- Study$new(
        ID = paste(studyId, comp$PartitionCoefficientMethod, comp$CellularPermeabilityMethod, sep = "_"), 
        compounds = list(comp), 
        individual = "Human"
      )
      
      # add to the list of studies
      Studies <- c(Studies, study)
    }
  }
  # add studies ID as names to the list of studies
  names(Studies) <- sapply(Studies, function(x) x$ID)
  
  return(Studies)
}
``` 

### Simulating the studies 

Once we have the studies created, we can run the simulation. Let' s run the studies for the first 
scenario.

```{r running_studies, eval = F}
studies <- createStudies(scenario = scenarios[1, ])
  
# Create folder to save the the pkml and results if wanted and run the simulations
outputFolderScenario <- file.path(outputFolder, "Chembl_Scenario_1")
dir.create(outputFolderScenario, showWarnings = FALSE)
  
# run Predictions for compound plasma in PVB
results <- runPredictions(
  studies,
  outputFolder = outputFolderScenario,
  numberOfCores = 5,
  outputSelections = c("Organism|PeripheralVenousBlood|**|Plasma (Peripheral Venous Blood)"),
  simulationResolution = c(0, max(InputStudies$`t_end [h]`) * 60, 0.1),
  saveResults = FALSE,
  saveSimulation = FALSE,
  queueSize = 200
)

```


### Calculating wanted prediction metrics

The output metrics have to be customized to the one required for a specific analysis here 
we have various AUC and Cmax and we will calculate them on our HT-PBPK predictions using `{PKNCA}` 
package to matched the observed dataset.

The metric part will depends on the observed data we need, so this part will also be adjusted to
one's need.

Given that we need to calculate the metrics for multiple studies and scenarios it could be easier 
to create a function that will do the calculations for us. 

```{r calculate_metrics_function, eval = TRUE}
calculateMetrics <- function(study, simData, obsData) {
  # if needed convert time unit to match obs data time unit in our case all time are in hours
  simData$data$Time <- ospsuite::toUnit(
    quantityOrDimension = simData$metaData$Time$dimension,
    values = simData$data$Time,
    sourceUnit = simData$metaData$Time$unit,
    targetUnit = "h"
  )
  simData$metaData$Time$unit <- "h"
  
  # add compound molecular weight to the metadata to ensure conversion of umol/l to g/l if needed
  simData$metaData$MW <- list(
    value = suppressMessages(study$Compounds[[1]]$getProperty("Molecular weight")$value), 
    unit = suppressMessages(study$Compounds[[1]]$getProperty("Molecular weight")$unit)
  )
 
  # Initialise result vector
  SimMetric <- c()
  
  # calculate metrics for each observed data (a single study can have multiple observed data)
  for (i in seq_len(nrow(obsData))) {  
    if (obsData[i,]$Metric %in% c("AUC", "Cmax")) {
      # Filter time required to calculate the AUC from the correct time interval (e.g. restrict 
      # to 0:24h for AUC024, restrict to 0 to end of sim for AUC0Inf, or to the last dose interval 
      # for Cmax of last dose, ...)
      simDataToUse <- simData$data %>% 
        dplyr::filter(
          Time <= min(obsData[i,]$`t_end [h]`, as.numeric(obsData[i,]$`Metric_t2 [h]`)), 
          Time >= obsData[i,]$`Metric_t1 [h]`, 
        )
    } else {
      # In our case the only other metric is Concentration at a specific timepoint
      simDataToUse <- simData$data
    } 
      
    if (obsData[i,]$Metric == "AUC") {
      # calculate AUC with PKNCA (suppress no dose given warning as not required for auc)
      resPred <- suppressWarnings(
        PKNCA::pk.nca(
          PKNCA::PKNCAdata(
            data.conc = PKNCA::PKNCAconc(
              simDataToUse, `Organism|PeripheralVenousBlood|Compound1|Plasma (Peripheral Venous Blood)` ~ Time|IndividualId
            ), 
            intervals = data.frame(
              start = obsData[i,]$`Metric_t1 [h]`, 
              end = as.numeric(obsData[i,]$`Metric_t2 [h]`), 
              auclast = TRUE,
              aucinf.pred = TRUE,
              aucinf.obs = FALSE,
              half.life = FALSE
            )
          )
        )
      )
      
      # get PKNCA results
      if (is.finite(as.numeric(obsData[i,]$`Metric_t2 [h]`))) {
        # for AUC t1 to finite time  
        resPred <- resPred$result %>% filter(PPTESTCD == "auclast") %>% pull(PPORRES)
      } else {
        # for AUC t1 to Inf 
        resPred <- resPred$result %>% filter(PPTESTCD == "aucinf.pred") %>% pull(PPORRES)
      }
      
      # convert to the observed unit
      resPred <- ospsuite::toUnit(
        quantityOrDimension = ospsuite::getDimensionForUnit(obsData[i,]$OspUnits),
        values = resPred,
        # add "*" time unit to the simulated concentration unit to unit to make compatible AUC units
        sourceUnit = gsub(
          pattern = "/", 
          replacement = paste0("*", simData$metaData$Time$unit, "/"),
          x = simData$metaData$`Organism|PeripheralVenousBlood|Compound1|Plasma (Peripheral Venous Blood)`$unit,
        ),
        targetUnit = obsData[i,]$OspUnits,
        molWeight = simData$metaData$MW$value,
        molWeightUnit = simData$metaData$MW$unit
      )
      
    } else if (obsData[i,]$Metric %in% c("Cmax", "C")) {
      # convert to same unit as observed data
      resPred <- ospsuite::toUnit(
        quantityOrDimension = simData$metaData$`Organism|PeripheralVenousBlood|Compound1|Plasma (Peripheral Venous Blood)`$dimension,
        values = simDataToUse$`Organism|PeripheralVenousBlood|Compound1|Plasma (Peripheral Venous Blood)`,
        sourceUnit = simData$metaData$`Organism|PeripheralVenousBlood|Compound1|Plasma (Peripheral Venous Blood)`$unit,
        targetUnit = obsData[i,]$OspUnits,
        molWeight = simData$metaData$MW$value,
        molWeightUnit = simData$metaData$MW$unit
      )
      
      if (obsData[i,]$Metric  == "Cmax") {
        resPred <- max(resPred)
      } else {
        # interpolate to get value at time
        resPred <- approx(
          x =  simData$data$Time, 
          y =  resPred, 
          xout = obsData[i,]$`Metric_t1 [h]`, 
          rule = 2
        )
        resPred <- resPred$y 
      }
    }
    SimMetric <- c(SimMetric, resPred)
  }
  
  return(cbind(obsData, SimMetric  = SimMetric))
}
```

Now that we created the function calculate the metrics on our simulated data, 
we can run it for each study results in parallel and save as a csv file. 

```{r calculate_metrics, eval = FALSE}
# create and register parallel backend for calculating the metric needed 
# to compare to the observed data faster
cl <- makePSOCKcluster(5)
registerDoParallel(cl)

# get predicted values from the HT-PBPK run first as it cannot be access from parallel workers
simData <- lapply(results, function(studyResult) {
  ospsuite::getOutputValues(studyResult, individualIds = results$allIndividualIds)
})

# Calculate observed metrics on the predicted curves in parallel
metric_results <- foreach::foreach(
  studyID = unique(InputStudies$StudyID), 
  .combine = rbind, 
  .packages = c("dplyr", "PKNCA", "ospsuite"),
  .export = c("InputStudies", "calculateMetrics", "simData", "studies"),
  .errorhandling = "pass",
  .verbose = T
) %dopar% {
  simID <- paste(studyID, "PK-Sim Standard", "PK-Sim Standard", sep = "_")

  calculateMetrics(
    study = studies[[simID]], 
    simData = simData[[simID]], 
    obsData = InputStudies %>% filter(StudyID == studyID)
  )
}  
  
write.csv(
  metric_results, 
  file = file.path(outputFolder, paste0("metrics_scenario_1.csv")), 
  row.names = FALSE
)

# stop cluster once done
stopCluster(cl)
``` 

### Testing multiple scenarios

As you can see it would be easy to run it for multiple scenario to test various QSARs or hypotheses.

```{r, eval = FALSE}
# create and register parallel backend for calculating the metric needed 
# to compare to the observed data faster
cl <- makePSOCKcluster(5)
registerDoParallel(cl)

# Create studies for each compound and each scenario we want (here we are testing default scenario
# and with increase organ permeability)

for (scenario_row in seq_len(nrow(scenarios))) {
  scenarioNb <- scenarios$Scenario[scenario_row]

  # create studies for the scenario
  studies <- createStudies(scenario = scenarios[scenario_row, ])
  
  # Create folder to save the the pkml and results if wanted and run the simulations
  outputFolderScenario <- file.path(outputFolder, paste0("Chembl_Scenario_", scenarioNb))
  dir.create(outputFolderScenario, showWarnings = FALSE)
  
  # run Predictions for compound plasma in PVB
  results <- runPredictions(
    studies,
    outputFolder = outputFolderScenario,
    numberOfCores = 5,
    outputSelections = c("Organism|PeripheralVenousBlood|**|Plasma (Peripheral Venous Blood)"),
    simulationResolution = c(0, max(InputStudies$`t_end [h]`) * 60, 0.1),
    saveResults = FALSE,
    saveSimulation = FALSE,
    queueSize = 200
  )
  
  # get predicted values from the HT-PBPK run first as it cannot be access from parallel workers
  simData <- lapply(results, function(studyResult) {
    ospsuite::getOutputValues(studyResult, individualIds = results$allIndividualIds)
  })

  # Calculate observed metrics on the predicted curves in parallel
  metric_results <- foreach::foreach(
    studyID = unique(InputStudies$StudyID), 
    .combine = rbind, 
    .packages = c("dplyr", "PKNCA", "ospsuite"),
    .export = c("InputStudies", "calculateMetrics", "simData", "studies"),
    .errorhandling = "pass",
    .verbose = T
  ) %dopar% {
    simID <- paste(studyID, "PK-Sim Standard", "PK-Sim Standard", sep = "_")
  
    calculateMetrics(
      study = studies[[simID]], 
      simData = simData[[simID]], 
      obsData = InputStudies %>% filter(StudyID == studyID)
    )
  }  
  
  write.csv(
    metric_results, 
    file = file.path(outputFolder, paste0("metrics_scenario", scenarioNb, ".csv")), 
    row.names = FALSE
  )
}

# stop cluster once done
stopCluster(cl)
```

## Looking at the results

### AUC0Inf for scenario 1

The results have already been calculated and are available in the package. 
You can load them using the following code. Here for the default scenario 1:

```{r load_results}
path <- system.file("extdata", "Chembl_Test_Pipeline", "metrics_scenario1.csv", package = "ESQhtpbpk")
metric_results_1 <- read.csv(path, stringsAsFactors = FALSE, check.names = FALSE)
``` 

The predicted can be check with observed data. 

The AUC 0-Inf for all compounds is centered slightly above 1, meaning with the default HT-PBPK and 
ADMETlab QSAR it tends to slightly overestimate the AUC compared to the observed value: 

```{r echo=FALSE, class.source = 'fold-hide'}
metric_results_1 %>% 
  filter(Metric == "AUC", `Metric_t1 [h]` == 0, `Metric_t2 [h]` == Inf) %>% 
  group_by(molecule_pref_name) %>%
  summarise(
    AUC0InfAvg = mean(Value, na.rm = TRUE), 
    AUC0InfPredAvg = mean(SimMetric, na.rm = TRUE),
    AUC0InfRatio = mean(SimMetric / Value, na.rm = TRUE)
  ) %>%
  ggplot() +
    geom_boxplot(aes(y = AUC0InfRatio)) + 
    scale_y_log10() + 
    theme_bw() + 
    theme(aspect.ratio = 5)
```

```{r include=FALSE}
tmp <- metric_results_1 %>% 
  filter(Metric == "AUC", `Metric_t1 [h]` == 0, `Metric_t2 [h]` == Inf) %>% 
  group_by(molecule_pref_name) %>%
  summarise(
    AUC0InfAvg = mean(Value, na.rm = TRUE), 
    AUC0InfPredAvg = mean(SimMetric, na.rm = TRUE),
    AUC0InfRatio = mean(SimMetric / Value, na.rm = TRUE)
  ) %>%
  summarise(
    AUC0InfRatio_2Fold = mean(AUC0InfRatio <= 2 & AUC0InfRatio >= 1/2, na.rm = TRUE) * 100,
    AUC0InfRatio_4Fold = mean(AUC0InfRatio <= 4 & AUC0InfRatio >= 1/4, na.rm = TRUE) * 100,
    AUC0InfRatio_10Fold = mean(AUC0InfRatio <= 10 & AUC0InfRatio >= 1/10, na.rm = TRUE) * 100,
  )
```

There are `r signif(tmp$AUC0InfRatio_2Fold,3)` % of compounds with a predicted AUC0Inf within 2-fold of the observed value, `r signif(tmp$AUC0InfRatio_4Fold, 3)`% within 4-fold and `r signif(tmp$AUC0InfRatio_10Fold,3)` % within 10-fold.

We can also look at the AUC0Inf fold ratio for the different percentiles:

```{r class.source = 'fold-hide'}
tmp <- metric_results_1 %>% 
  filter(Metric == "AUC", `Metric_t1 [h]` == 0, `Metric_t2 [h]` == Inf) %>% 
  group_by(molecule_pref_name) %>%
  summarise(
    AUC0InfAvg = mean(Value, na.rm = TRUE), 
    AUC0InfPredAvg = mean(SimMetric, na.rm = TRUE),
    AUC0InfRatio = mean(SimMetric / Value, na.rm = TRUE)
  ) %>%
  summarise(
    AUC0InfRatio5th = 2^quantile(abs(log2(AUC0InfRatio)), c(0.05)),
    AUC0InfRatio25th = 2^quantile(abs(log2(AUC0InfRatio)), c(0.25)),
    AUC0InfRatio50th = 2^quantile(abs(log2(AUC0InfRatio)), c(0.50)),
    AUC0InfRatio75th = 2^quantile(abs(log2(AUC0InfRatio)), c(0.75)),
    AUC0InfRatio95th = 2^quantile(abs(log2(AUC0InfRatio)), c(0.95))
  )

tmp
```
So We have 50% of the compound for which the AUC0Inf fold ratio is within `r tmp$AUC0InfRatio50th` folds.

If we look more in detail at the AUC 0-Inf 
(one dot is one studies and one compound i.e. one color can have multiple studies associated), 
we can see that most of the studies/compounds are within 10-fold of the observed value, with one 
compound being particularly underpredicted.

```{r class.source = 'fold-hide'}
p <- metric_results_1 %>% filter(Metric == "AUC", `Metric_t1 [h]` == 0, `Metric_t2 [h]` == Inf) %>% 
  ggplot(aes(x = Value, y = SimMetric, color = molecule_pref_name)) + 
  geom_point(show.legend = F) + 
  geom_abline(slope = 1, intercept = 0) + 
  labs(x = "Observed AUC0Inf", y = "Predicted AUC0Inf") + 
  scale_x_log10() + scale_y_log10() +
  geom_abline(slope = 1, intercept = log10(2), col = "grey50", linetype = "dashed") + 
  geom_abline(slope = 1, intercept = -log10(2), col = "grey50", linetype = "dashed") + 
  geom_abline(slope = 1, intercept = log10(10), col = "grey80", linetype = "dashed") + 
  geom_abline(slope = 1, intercept = -log(10), col = "grey80", linetype = "dashed") + 
  theme_bw()

ggplotly(p)
```

### Cmax for scenario 1

Similarly we can also look at other metrics such as Cmax. 
The Cmax for all compounds tend to be overpredicted.

```{r echo=FALSE, class.source = 'fold-hide'}
metric_results_1 %>% 
  filter(Metric == "Cmax", `Metric_t1 [h]` == 0) %>% 
  group_by(molecule_pref_name) %>%
  summarise(
    CmaxAvg = mean(Value, na.rm = TRUE), 
    CmaxPredAvg = mean(SimMetric, na.rm = TRUE),
    CmaxRatio = mean(SimMetric / Value, na.rm = TRUE)
  ) %>%
  ggplot() +
    geom_boxplot(aes(y = CmaxRatio)) + 
    scale_y_log10() + 
    theme_bw() + 
    theme(aspect.ratio = 5)
```

```{r include=FALSE, class.source = 'fold-hide'}
tmp <- metric_results_1 %>% 
  filter(Metric == "Cmax", `Metric_t1 [h]` == 0) %>% 
  group_by(molecule_pref_name) %>%
  summarise(
    CmaxAvg = mean(Value, na.rm = TRUE), 
    CmaxPredAvg = mean(SimMetric, na.rm = TRUE),
    CmaxRatio = mean(SimMetric / Value, na.rm = TRUE)
  ) %>%
  summarise(
    CmaxRatio_2Fold = mean(CmaxRatio <= 2 & CmaxRatio >= 1/2, na.rm = TRUE) * 100,
    CmaxRatio_4Fold = mean(CmaxRatio <= 4 & CmaxRatio >= 1/4, na.rm = TRUE) * 100,
    CmaxRatio_10Fold = mean(CmaxRatio <= 10 & CmaxRatio >= 1/10, na.rm = TRUE) * 100,
  )
```

There are `r signif(tmp$CmaxRatio_2Fold,3)` % of compounds with a predicted Cmax within 2-fold of the observed value, `r signif(tmp$CmaxRatio_4Fold, 3)`% within 4-fold and `r signif(tmp$CmaxRatio_10Fold,3)` % within 10-fold.

We can also look at the AUC0Inf fold ratio for the different percentiles:

```{r class.source='fold-hide'}
metric_results_1 %>% 
  filter(Metric == "Cmax", `Metric_t1 [h]` == 0) %>% 
  group_by(molecule_pref_name) %>%
  summarise(
    CmaxAvg = mean(Value, na.rm = TRUE), 
    CmaxPredAvg = mean(SimMetric, na.rm = TRUE),
    CmaxRatio = mean(SimMetric / Value, na.rm = TRUE)
  ) %>%
  summarise(
    CmaxRatio5th = 2^quantile(abs(log2(CmaxRatio)), c(0.05)),
    CmaxRatio25th = 2^quantile(abs(log2(CmaxRatio)), c(0.25)),
    CmaxRatio50th = 2^quantile(abs(log2(CmaxRatio)), c(0.50)),
    CmaxRatio75th = 2^quantile(abs(log2(CmaxRatio)), c(0.75)),
    CmaxRatio95th = 2^quantile(abs(log2(CmaxRatio)), c(0.95))
  )
```
So We have 50% of the compound for which the AUC0Inf fold ratio is within `r tmp$CmaxRatio50th` folds.

The comparison of Cmax can be slightly more difficult as the sample time points could affect the observed estimated value, 
and the full observed pk profiles are not available in the extracted Chembl data. So it could be expected that
the simulated Cmax with was sampled every 10 minutes could be different than the observed one, for which we 
we don't necessarily now how often the data were sampled or even how the Cmax  was derived (from maximal value 
observed or interpolated from NCA or 1-2-3 compartmental analysis, ...).

If we look more in detail at the Cmax 
(one dot is one studies and one compound i.e. one color can have multiple studies associated), 
we can see that most of the studies/compounds are overpredicted but within 10-fold of the observed value, with one 
compound being particularly underpredicted.

```{r, class.source = 'fold-hide'}
p <- metric_results_1 %>% filter(Metric == "Cmax", `Metric_t1 [h]` == 0) %>% 
  ggplot(aes(x = Value, y = SimMetric, color = molecule_pref_name)) + 
  geom_point(show.legend = F) + 
  geom_abline(slope = 1, intercept = 0) + 
  labs(x = "Observed Cmax", y = "Predicted Cmax") + 
  scale_x_log10() + scale_y_log10() +
  geom_abline(slope = 1, intercept = log10(2), col = "grey50", linetype = "dashed") + 
  geom_abline(slope = 1, intercept = -log10(2), col = "grey50", linetype = "dashed") + 
  geom_abline(slope = 1, intercept = 1, col = "grey80", linetype = "dashed") + 
  geom_abline(slope = 1, intercept = -1, col = "grey80", linetype = "dashed") + 
  theme_bw()

ggplotly(p)

```
### AUC vs Cmax for scenario 1

Looking at AUC and Cmax ratio together:  
```{r, class.source = "fold-hide"}
p <- metric_results_1 %>% 
  filter((Metric == "Cmax" & `Metric_t1 [h]` == 0 ) | (Metric == "AUC" & `Metric_t1 [h]` == 0 & `Metric_t2 [h]` == Inf)) %>% 
  tidyr::pivot_wider(
    id_cols = c(StudyID, ProtocolID, molecule_pref_name, canonical_smiles, Meal), 
    names_from = Metric, 
    values_from = c(Value, SimMetric), 
    values_fn = c("Value" = ~ mean(.x), "SimMetric" = ~ mean(.x))
  ) %>%
  mutate(AUC0Inf_Ratio = SimMetric_AUC / Value_AUC, 
         Cmax_Ratio = SimMetric_Cmax / Value_Cmax) %>%
  ggplot(aes(x = AUC0Inf_Ratio, y = Cmax_Ratio, color = molecule_pref_name)) + 
  geom_point(show.legend = F) + 
  geom_abline(slope = 1, intercept = 0) + 
  labs(x = "AUC0Inf ratio", y = "Cmax ratio") + 
  scale_x_log10() + scale_y_log10() +
  geom_vline(xintercept = 2, col = "grey50", linetype = "dashed") + 
  geom_vline(xintercept = 0.5, col = "grey50", linetype = "dashed") + 
  geom_vline(xintercept = 10, col = "grey80", linetype = "dashed") + 
  geom_vline(xintercept = 0.1, col = "grey80", linetype = "dashed") + 
  geom_hline(yintercept = 2, col = "grey50", linetype = "dashed") + 
  geom_hline(yintercept = 0.5, col = "grey50", linetype = "dashed") + 
  geom_hline(yintercept = 10, col = "grey80", linetype = "dashed") + 
  geom_hline(yintercept = 0.1, col = "grey80", linetype = "dashed") + 
  theme_bw()

ggplotly(p)

```


### Characteristic of compound compared to their predictability

We can additional try to understand for which compounds the AUC and Cmax are less well predicted, 
to better understand the potential applicability domain of the pipeline even though
it will be mostly constrained by applicability domain of the used QSAR inputs as well as the PBPK 
model. 

#### Lipophilicity effect

Compounds with high or low lipophilicty values tends to be the 
most problematic ones and not well predicted. 
 
```{r, class.source = 'fold-hide'}

p <- full_join(
  x = metric_results_1 %>% 
    filter(Metric == "AUC" & `Metric_t1 [h]` == 0 & `Metric_t2 [h]` == Inf) %>% 
    mutate(AUC0Inf_Ratio = SimMetric / Value),
  y = InputCompounds, 
  by = c("molecule_pref_name", "canonical_smiles")
) %>% 
  ggplot(aes(x = logD, y = AUC0Inf_Ratio, color = molecule_pref_name)) + 
  geom_point() + 
  geom_hline(yintercept = 2, col = "grey50", linetype = "dashed") + 
  geom_hline(yintercept = 0.5, col = "grey50", linetype = "dashed") + 
  geom_hline(yintercept = 10, col = "grey80", linetype = "dashed") + 
  geom_hline(yintercept = 0.1, col = "grey80", linetype = "dashed") + 
  scale_y_log10() + 
  labs(x = "Compound logD", y = "AUC0Inf ratio") + 
  theme_bw() 

ggplotly(p)
```

We observed a trend that high lypophilic compounds have an underpredicted AUC0Inf, while low lipophilicity 
compounds are more overpredicted. 

#### Fraction unbound effect

Similarly compounds that have a very low fraction unbound are 
also poorly predicted.

```{r, class.source = 'fold-hide'}
p <- full_join(
  x = metric_results_1 %>% 
    filter(Metric == "AUC" & `Metric_t1 [h]` == 0 & `Metric_t2 [h]` == Inf) %>% 
    mutate(AUC0Inf_Ratio = SimMetric / Value),
  y = InputCompounds, 
  by = c("molecule_pref_name", "canonical_smiles")
) %>% 
  ggplot(aes(x = Fu, y = AUC0Inf_Ratio, color = molecule_pref_name)) + 
  geom_point() + 
  geom_hline(yintercept = 2, col = "grey50", linetype = "dashed") + 
  geom_hline(yintercept = 0.5, col = "grey50", linetype = "dashed") + 
  geom_hline(yintercept = 10, col = "grey80", linetype = "dashed") + 
  geom_hline(yintercept = 0.1, col = "grey80", linetype = "dashed") + 
  scale_y_log10() + 
  labs(x = "Compound Fraction unbound", y = "AUC0Inf ratio") + 
  theme_bw() 

ggplotly(p)

```

#### Solubility effect

Similarly compounds that have a very low solubility are 
also poorly predicted.

```{r class.source = 'fold-hide'}
p <- full_join(
  x = metric_results_1 %>% 
    filter(Metric == "AUC" & `Metric_t1 [h]` == 0 & `Metric_t2 [h]` == Inf) %>% 
    mutate(AUC0Inf_Ratio = SimMetric / Value),
  y = InputCompounds, 
  by = c("molecule_pref_name", "canonical_smiles")
) %>% 
  ggplot(aes(x = logS, y = AUC0Inf_Ratio, color = molecule_pref_name)) + 
  geom_point() + 
  geom_hline(yintercept = 2, col = "grey50", linetype = "dashed") + 
  geom_hline(yintercept = 0.5, col = "grey50", linetype = "dashed") + 
  geom_hline(yintercept = 10, col = "grey80", linetype = "dashed") + 
  geom_hline(yintercept = 0.1, col = "grey80", linetype = "dashed") + 
  scale_y_log10() + 
  labs(x = "Compound Solubility (logS)", y = "AUC0Inf ratio") + 
  theme_bw() 

ggplotly(p)
```

### Compare with other scenario

We can compare with other scenarios that use different QSARs as input make different assumptions (
for example adding artificially high organ permeability (scenario 09))

```{r, load_various_scenarios}
for (i in 2:15) {
  path <- system.file("extdata", "Chembl_Test_Pipeline", paste0("metrics_scenario", i, ".csv"), package = "ESQhtpbpk")
  assign(x = paste0("metric_results_", i), value = read.csv(path, stringsAsFactors = FALSE, check.names = FALSE))
}
``` 

Calculated and compare AUC 0 Inf ratio and Cmax of the different scenarios:

```{r class.source = "fold-hide"}
scenario_1to15 <- metric_results_1 %>% mutate(Scenario = "Scenario 1")

for (i in 2:15) {
  scenario_1to15 <- rbind(
    scenario_1to15, 
    eval(parse(text = paste0("metric_results_", i))) %>% 
      mutate(Scenario = paste("Scenario", i))
  )
}
```

```{r class.source = "fold-hide"}
tmp <- scenario_1to15 %>% filter(Metric == "AUC", `Metric_t1 [h]` == 0, `Metric_t2 [h]` == Inf) %>%
  group_by(molecule_pref_name, Scenario) %>%
  summarise(
    AUC0InfAvg = mean(Value, na.rm = TRUE), 
    AUC0InfPredAvg = mean(SimMetric, na.rm = TRUE),
    AUC0InfRatio = mean(SimMetric / Value, na.rm = TRUE)
  )
tmp$Scenario <- factor(tmp$Scenario, levels = paste("Scenario", 1:15), ordered = TRUE)

p <- ggplot(tmp, aes(y = AUC0InfRatio, x = Scenario)) + 
  geom_boxplot() +
  scale_y_log10() + 
  theme_bw() + 
  geom_hline(yintercept = 2, col = "grey50", linetype = "dashed") +
  geom_hline(yintercept = 0.5, col = "grey50", linetype = "dashed") +
  geom_hline(yintercept = 10, col = "grey20", linetype = "dashed") +
  geom_hline(yintercept = 0.1, col = "grey20", linetype = "dashed") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p
```


```{r class.source = "fold-hide"}
tmp <- scenario_1to15 %>% filter(Metric == "Cmax", `Metric_t1 [h]` == 0) %>%
  group_by(molecule_pref_name, Scenario) %>%
  summarise(
    CmaxAvg = mean(Value, na.rm = TRUE), 
    CmaxPredAvg = mean(SimMetric, na.rm = TRUE),
    CmaxRatio = mean(SimMetric / Value, na.rm = TRUE)
  )
tmp$Scenario <- factor(tmp$Scenario, levels = paste("Scenario", 1:15), ordered = TRUE)

p <- ggplot(tmp, aes(y = CmaxRatio, x = Scenario)) + 
  geom_boxplot() +
  scale_y_log10() + 
  theme_bw() + 
  geom_hline(yintercept = 2, col = "grey50", linetype = "dashed") +
  geom_hline(yintercept = 0.5, col = "grey50", linetype = "dashed") +
  geom_hline(yintercept = 10, col = "grey20", linetype = "dashed") +
  geom_hline(yintercept = 0.1, col = "grey20", linetype = "dashed") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  coord_cartesian(ylim = c(0.001, 1000))
p
```

Scenario 9 (assuming high organ permeability) and 15 (using logMA as calculated by Pearce for Lipophilicity) 
seems to give more reasonable AUC0Inf and Cmax. Even for Scenario 15 still tends to be overpredicted in most cases

We will those more in detail.
