---
title: "Example with chembl dataset"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

For this example we will load a few libraries

```{r setup}
library(ESQhtpbpk) # package for HTPBPK
library(readxl) # package to read xlsx files that contains input data
library(dplyr) # package for easy manipulation of data
library(tidyr) # package for easy manipulation of data
library(stringr) # package for string manipulation
# packages for plotting
library(ggplot2)
library(ggrepel)
library(scales)
# packages for parallel computing
library(doParallel)
```

Based on [this paper](https://link.springer.com/article/10.1007/s00204-024-03764-9) and more specifically [Table 1](https://link.springer.com/article/10.1007/s00204-024-03764-9/tables/1), we will use QSAR from ADMETlab as input to the HT-PBPK model. A dataset of Cmax and AUC extracted and curated from CHEMBL is available in the package as an example dataset.

```{r loading_datasets}
# Reading curated chembl dataset with AUC and Cmax values
path <- system.file("extdata", "Chembl_preprocessed_curated.xlsx", package = "ESQhtpbpk")

InputStudies <- read_excel(path, sheet = 1)
InputProtocol <- read_excel(path, sheet = 2)
InputCompounds <- read_excel(path, sheet = 3) # QSARs obtained from ADMETlab 3.0
```

The platform allows to test easily multiple scenarios by adjusting inputs as needed based on the chosen scenarios. 
An example of excel file with different scenarios is available in the package.

```{r load_scenarios}
scenarios <- read.csv(
  file = system.file("extdata", "Chembl_Test_Pipeline", "scenarios.csv", package = "ESQhtpbpk"), 
  stringsAsFactors = FALSE
)

head(scenarios)
```

The output metrics have to be customized to the one required for a specific analysis here various AUC and Cmax are calculated to matched the observed dataset.

```{r multiple scenarios, eval = F}
outputFolder <- "Chembl_Test_Pipeline"
dir.create(outputFolder, showWarnings = FALSE)


# register  parallel  backend 
cl <- makePSOCKcluster(5, outfile = file.path(outputFolder, "parallel_log.txt"))
registerDoParallel(cl)

# Create studies for each compound and each scenario
for (scenario_number in seq_len(nrow(scenarios))) {
  Studies <- list()

  for (compoundIdx in seq_len(nrow(InputCompounds))) {
    compName <- InputCompounds$molecule_chembl_id[compoundIdx]
    
    # set new compound with compound name
    comp <- Compound$new(ID = compName)
  
    # if halogens are considered set halogenes numbers
    if (scenarios$Halogen[scenario_number] == 1) {
      # Set halogen numbers p to 10 allowed in PK-Sim
      comp$setPropertyValue("Chlorine count", value = min(10, str_count(InputCompounds$smiles[compoundIdx], "Cl")))
      comp$setPropertyValue("Bromine count", value = min(10, str_count(InputCompounds$smiles[compoundIdx], "Br")))
      comp$setPropertyValue("Fluorine count", value = min(10, str_count(InputCompounds$smiles[compoundIdx], "F")))
      comp$setPropertyValue("Iodine count", value = min(10, str_count(InputCompounds$smiles[compoundIdx], "I")))
    }
    
    # set molecular weight
    comp$setPropertyValue("Molecular weight", InputCompounds$MW[compoundIdx], unit = "g/mol")
    
    # set solubity depending on scenario
    switch(
      as.character(scenarios$Solubility[scenario_number]), 
      logS = {
        # set solubility in g/l
        comp$setPropertyValue("Solubility", 10^InputCompounds$logS[compoundIdx] * InputCompounds$MW[compoundIdx], unit = "g/l") # logS is in mol/l
      }, 
      High = {
        # set arbitrary high solubility
        comp$setPropertyValue("Solubility", 1000, unit = "g/l")
      }
    )
    
    # set lipophilicity based on scenario
    switch(
      as.character(scenarios$Lipophilicity[scenario_number]), 
      logP = {
        # set logP
        comp$setPropertyValue("Lipophilicity", InputCompounds$logP[compoundIdx], unit = "Log Units")
      },
      logD = {
        # use logD as lipophilicity input
        comp$setPropertyValue("Lipophilicity", InputCompounds$logD[compoundIdx], unit = "Log Units")
      },
      logMA1 = {
        # use logMA as lipophilicity input
         comp$setPropertyValue("Lipophilicity", 1.294 + 0.304 * InputCompounds$logP[compoundIdx], unit = "Log Units")
      },
      logMA_Pearce = {
        # use logMA as lipophilicity input
        comp$setPropertyValue("Lipophilicity", 0.0166 * 37 + 0.882 * InputCompounds$logP[compoundIdx], unit = "Log Units")
      }
    )
    
    # set Fu based on scenario
    switch(
      as.character(scenarios$Fu[scenario_number]), 
      Fu = {
        comp$setPropertyValue("Fraction unbound", InputCompounds$Fu[compoundIdx], unit = "%")
      },
      PPB = {
        comp$setPropertyValue("Fraction unbound", 100 - InputCompounds$PPB[compoundIdx], unit = "%")
      }
    )

    # Set intestinal permeability based on scenario
    switch(
      as.character(scenarios$Pint[scenario_number]), 
      MDCK = {
        # set Pint to MDCK value
        comp$addProperty(
          name = "PInt",
          parName = "Specific intestinal permeability (transcellular)",
          dimension = "Velocity",
          value = 10^InputCompounds$MDCK[compoundIdx], # MDCK is in log cm/s,
          unit = "cm/s"
        )
        comp$setPropertyValue("Fraction unbound", InputCompounds$Fu[compoundIdx], unit = "%")
      },
      Caco2 = {
        # set Pint to Caco2 value
        comp$addProperty(
          name = "PInt",
          parName = "Specific intestinal permeability (transcellular)",
          dimension = "Velocity",
          value = 10^InputCompounds$caco2[compoundIdx], # Caco2 is in log cm/s,
          unit = "cm/s"
        )
      },
      PAMPA = {
        # set Pint to PAMPA value
        comp$addProperty(
          name = "PInt",
          parName = "Specific intestinal permeability (transcellular)",
          dimension = "Velocity",
          value = InputCompounds$PAMPA[compoundIdx] * 1e-6, # PAMPA is in 10E-6 cm/s,
          unit = "cm/s"
        )
      }
    )
    
    # set Perm based on scenario arbitrarly high
    if (scenarios$Perm[scenario_number] ==  "High") {
      comp$addProperty(
        name = "Perm",
        parName = "Permeability",
        dimension = "Velocity",
        value = 10,
        unit = "cm/s"
      )
    }
   
    # add total clearance (as hepatic clearance)
    comp$addProcessProperty(
      processType = "Liver Plasma Clearance",
      propertyName = "Plasma clearance",
      parName = "Plasma clearance",
      dimension = "Flow per weight",
      value = InputCompounds$`cl-plasma`[compoundIdx], 
      unit = "ml/min/kg"
    )
    comp$addProcessProperty(
      processType = "Liver Plasma Clearance",
      propertyName = "Lipophilicity",
      parName = "Lipophilicity (experiment)",
      dimension = "Log Units",
      value = suppressMessages(comp$getProperty("Lipophilicity")$value),
      unit = suppressMessages(comp$getProperty("Lipophilicity")$unit)
    )
    comp$addProcessProperty(
      processType = "Liver Plasma Clearance",
      propertyName = "Fraction unbound",
      parName = "Fraction unbound (experiment)",
      dimension = "Fraction",
      value = suppressMessages(comp$getProperty("Fraction unbound")$value),
      unit = suppressMessages(comp$getProperty("Fraction unbound")$unit)
    )
    
    if (scenarios$PkA[scenario_number] == 1) {
      comp$setPropertyValue(
        name = "pKa value 0", 
        value = min(max(c(0,InputCompounds$pka_acidic[compoundIdx])), 14)
        )
      comp$setPropertyValue("Compound type 0", "Acidic")
      
      comp$setPropertyValue(
        name = "pKa value 1", 
        value = min(max(c(0,InputCompounds$pka_basic[compoundIdx])), 14)
      )
      comp$setPropertyValue("Compound type 1", "Basic")
    }
    
    if (scenarios$GFR[scenario_number] == 1) {
      comp$addProcessProperty(
        processType = "GFR",
        propertyName = "GFR",
        parName = "GFR fraction",
        dimension = "Dimensionless",
        value = 1
      )
    }

    if (scenarios$Enzyme_correction[scenario_number] == 1) {
      # set enzyme correction
      comp$addProcessProperty(
        processType = "Liver Plasma Clearance",
        propertyName = "Enzyme correction",
        parName = "Enzyme concentration",
        dimension = "Concentration (molar)",
        value = 1/0.667, # 1/0.667 is the correction factor for the enzyme activity
        unit = "umol/l"
      )
    }
    
    # create studies using the compound
    for (studyId in unique(InputStudies$StudyID[which(InputStudies$molecule_chembl_id == compName)])) {
      print(studyId)
      studyIdx <- which(InputStudies$StudyID == studyId)
      studyProtocolId <- unique(InputStudies$ProtocolID[studyIdx])
      protocolTable <- InputProtocol[InputProtocol$ProtocolID == studyProtocolId,]
      
      # create advance protocol (1 protocol can be defie by multiple line in excel with different 
      # admin, for example with loading dose)
      prot <- AdvancedProtocol$new()
      
      # skip is dose norm not available
      if (any(is.na(protocolTable$`Dose Norm`))) {
        next()
      }
      
      for (i in seq_len(nrow(protocolTable))) {
        prot$addSchema(
          schemaName = paste0("Schema_", i),
          startTime = protocolTable$Time[i],
          timeBetweenRepetitions = ifelse(is.na(protocolTable$Time_between_rep[i]), 0, protocolTable$Time_between_rep[i]),
          numberOfRepetitions = ifelse(is.na(protocolTable$Nrep[i]), 1, protocolTable$Nrep[i]),
          timeUnit = protocolTable$TimeUnit[i]
        )
        
        switch(
          as.character(scenarios$Dose[scenario_number]), 
          Flat = {
            # set dose as flat dose (or BW if given as BW)
            admin <- SimpleProtocol$new(
              route = c("ORAL" = "Oral", "IV INFUSION" = "IV Infusion")[protocolTable$Route[i]],
              dose = protocolTable$`Dose`[i],
              doseUnit = protocolTable$`Dose Unit`[i],
              infusionTime = if (is.na(protocolTable$Infusion_duration[i])) {NULL} else {protocolTable$Infusion_duration[i]},
              infusionTimeUnit = if (is.na(protocolTable$Infusion_duration_unit[i])) {
                  NULL
                } else { 
                  protocolTable$Infusion_duration_unit[i] 
                }
            )
          },
          BW = {
            admin <- SimpleProtocol$new(
              route = c("ORAL" = "Oral", "IV INFUSION" = "IV Infusion")[protocolTable$Route[i]],
              dose = protocolTable$`Dose Norm`[i],
              doseUnit = protocolTable$`Dose Norm Unit`[i],
              infusionTime = if (is.na(protocolTable$Infusion_duration[i])) {NULL} else {protocolTable$Infusion_duration[i]},
              infusionTimeUnit = if (is.na(protocolTable$Infusion_duration_unit[i])) {
                  NULL
                } else { 
                  protocolTable$Infusion_duration_unit[i] 
                }
            )
          }
        )
        
        if (protocolTable$Route[i] == "Oral") {
          formulation <- createDissolvedFormulation(name = "Dissolved")
          admin$setFormulation(formulation = formulation)
        }
        
        prot$addProtocolToSchema(admin, schemaName = paste0("Schema_", i))
      }
      comp$setProtocol(prot)
      
      comp$PartitionCoefficientMethod <- "PK-Sim"
      comp$CellularPermeabilityMethod <- "PK-Sim"
          
      study <- Study$new(ID = paste(studyId, comp$PartitionCoefficientMethod, comp$CellularPermeabilityMethod, sep = "_"), 
                         compounds = list(comp), individual = "Human")
          
      Studies <- c(Studies, study)
    }
  }
  names(Studies) <- sapply(Studies, function(x) x$ID)
  
  
  # Create folder to save the results and run the simulations
  outputFolderScenario <- file.path(outputFolder, paste0("Chembl_Scenario_", scenarios$Scenario[scenario_number]))
  dir.create(outputFolderScenario, showWarnings = FALSE)
  results <- runPredictions(
    Studies,
    outputFolder = outputFolderScenario,
    numberOfCores = 5,
    outputSelections = c("Organism|PeripheralVenousBlood|**|Plasma (Peripheral Venous Blood)"),
    simulationResolution = c(0, max(InputStudies$`t_end [h]`) * 60, 0.1),
    saveResults = FALSE,
    saveSimulation = FALSE,
    queueSize = 200
  )
  
  predData <- list()
  #  get predicted data 
  for (studyID in unique(InputStudies$StudyID)) {

    # extract sim data 
    simID <- paste(studyID, "PK-Sim Standard", "PK-Sim Standard", sep = "_")
    predDataTmp <- ospsuite::getOutputValues(results[[simID]], individualIds = results$allIndividualIds)
    
    # convert time unit to match obs data
    predDataTmp$data$Time <- ospsuite::toUnit(
        quantityOrDimension = predDataTmp$metaData$Time$dimension,
        values = predDataTmp$data$Time,
        sourceUnit = predDataTmp$metaData$Time$unit,
        targetUnit = "h"
      )
    predDataTmp$metaData$Time$unit <- "h"
    predDataTmp$metaData$MW <- list(
      value = suppressMessages(Studies[[simID]]$Compounds[[1]]$getProperty("Molecular weight")$value), 
      unit = suppressMessages(Studies[[simID]]$Compounds[[1]]$getProperty("Molecular weight")$unit)
    )
                                                            
    
    predData <- c(predData, list(predDataTmp))
    names(predData)[length(predData)] <- studyID
  }
  
  print("Calculating metrics")
  # Calculate observed metrics on the predicted curves
  metric_results <- foreach::foreach(
    i = seq_len(nrow(InputStudies)), 
    .combine = rbind, 
    .packages = c("dplyr", "PKNCA", "ospsuite"),
    .export = c("predData", "InputStudies"),
    .errorhandling = "pass",
    .verbose = T
  ) %dopar% {
    studyID <- InputStudies$StudyID[i]
    obsData <- InputStudies[i,]

    print(i)
    if (obsData$Metric %in% c("AUC", "Cmax")) {
      simData <- predData[[studyID]]$data %>% 
        dplyr::filter(
          Time <= min(obsData$`t_end [h]`, as.numeric(obsData$`Metric_t2 [h]`)), 
          Time >= obsData$`Metric_t1 [h]`, 
      )
    } else {
      simData <- predData[[studyID]]$data
    } 
      
    if (obsData$Metric == "AUC") {
      resPred <- PKNCA::pk.nca(
        PKNCA::PKNCAdata(
          data.conc = PKNCA::PKNCAconc(
            simData, `Organism|PeripheralVenousBlood|Compound1|Plasma (Peripheral Venous Blood)`~Time|IndividualId
          ), 
          data.dose = PKNCA::PKNCAdose(
            data.frame(IndividualId = 0, dose = 1), dose~1|IndividualId
          ), 
          intervals = data.frame(
            start = obsData$`Metric_t1 [h]`, 
            end = as.numeric(obsData$`Metric_t2 [h]`), 
            auclast = TRUE,
            aucinf.pred = TRUE,
            aucinf.obs = TRUE,
            half.life = FALSE
          )
        )
      )
        
      if (is.finite(as.numeric(obsData$`Metric_t2 [h]`))) {
        resPred <- resPred$result %>% filter(PPTESTCD == "aucinf.pred") %>% pull(PPORRES)
      }else {
        resPred <- resPred$result %>% filter(PPTESTCD == "auclast") %>% pull(PPORRES)
      }
      
      resPred <- ospsuite::toUnit(
        quantityOrDimension = "AUC (mass)",
        values = resPred,
        sourceUnit = gsub(
          pattern = "/", 
          replacement = paste0("*", predData[[studyID]]$metaData$Time$unit, "/"),
          x = predData[[studyID]]$metaData$`Organism|PeripheralVenousBlood|Compound1|Plasma (Peripheral Venous Blood)`$unit,
        ),
        targetUnit = obsData$OspUnits,
        molWeight = predData[[studyID]]$metaData$MW$value,
        molWeightUnit = predData[[studyID]]$metaData$MW$unit
      )
      
    } else if (obsData$Metric %in% c("Cmax", "C")) {
      resPred <- ospsuite::toUnit(
        quantityOrDimension = predData[[studyID]]$metaData$`Organism|PeripheralVenousBlood|Compound1|Plasma (Peripheral Venous Blood)`$dimension,
        values = simData$`Organism|PeripheralVenousBlood|Compound1|Plasma (Peripheral Venous Blood)`,
        sourceUnit = predData[[studyID]]$metaData$`Organism|PeripheralVenousBlood|Compound1|Plasma (Peripheral Venous Blood)`$unit,
        targetUnit = obsData$OspUnits,
        molWeight = predData[[studyID]]$metaData$MW$value,
        molWeightUnit = predData[[studyID]]$metaData$MW$unit
      )
      
      if (obsData$Metric  == "Cmax") {
        resPred <- max(resPred)
      } else {
        # interpolate to get value at time
        resPred <- approx(
          x =  predData[[studyID]]$data$Time, 
          y =  resPred, 
          xout = obsData$`Metric_t1 [h]`, 
          rule = 2
        )
        resPred <- resPred$y 
      }
    }
    return( cbind(obsData, SimMetric  = resPred))
  }
  
  write.csv(
    metric_results, 
    file = file.path(outputFolder, paste0("metrics_", "scenario_", scenarios$Scenario[scenario_number], ".csv")), 
    row.names = FALSE
  )
} 

stopCluster(cl)
```

The results have already been calculated and are available in the package. 
You can load them using the following code. Here for the default scenario 1:

```{r load_results}
path <- system.file("extdata", "Chembl_Test_Pipeline", "metric_scenario_1.csv", package = "ESQhtpbpk")
metric_results_01 <- read.csv(path, stringsAsFactors = FALSE, check.names = FALSE)
``` 

The predicted can be check with observed data. 

The AUC 0-Inf for all compounds is centered around 1 the observed value: 
```{r echo=FALSE}
metric_results_01 %>% 
  filter(Metric == "AUC", `Metric_t1 [h]` == 0, `Metric_t2 [h]` == Inf) %>% 
  group_by(molecule_pref_name) %>%
  summarise(
    AUC0InfAvg = mean(Value, na.rm = TRUE), 
    AUC0InfPredAvg = mean(SimMetric, na.rm = TRUE),
    AUC0InfRatio = mean(SimMetric / Value, na.rm = TRUE)
  ) %>%
  ggplot() +
    geom_boxplot(aes(y = AUC0InfRatio)) + 
    scale_y_log10() + 
    theme_bw() + 
    theme(aspect.ratio = 5)
```

```{r include=FALSE}
tmp <- metric_results_01 %>% 
  filter(Metric == "AUC", `Metric_t1 [h]` == 0, `Metric_t2 [h]` == Inf) %>% 
  group_by(molecule_pref_name) %>%
  summarise(
    AUC0InfAvg = mean(Value, na.rm = TRUE), 
    AUC0InfPredAvg = mean(SimMetric, na.rm = TRUE),
    AUC0InfRatio = mean(SimMetric / Value, na.rm = TRUE)
  ) %>%
  summarise(
    AUC0InfRatio_2Fold = mean(AUC0InfRatio <= 2 & AUC0InfRatio >= 1/2, na.rm = TRUE) * 100,
    AUC0InfRatio_4Fold = mean(AUC0InfRatio <= 4 & AUC0InfRatio >= 1/4, na.rm = TRUE) * 100,
    AUC0InfRatio_10Fold = mean(AUC0InfRatio <= 10 & AUC0InfRatio >= 1/10, na.rm = TRUE) * 100,
  )
```
There are `r signif(tmp$AUC0InfRatio_2Fold,3` % of compounds with a predicted AUC0Inf within 2-fold of the observed value, `r signif(tmp$AUC0InfRatio_4Fold, 3)`% within 4-fold and `r signif(tmp$AUC0InfRatio_10Fold,3)` % within 10-fold.

```{r}
metric_results_01 %>% 
  filter(Metric == "AUC", `Metric_t1 [h]` == 0, `Metric_t2 [h]` == Inf) %>% 
  group_by(molecule_pref_name) %>%
  summarise(
    AUC0InfAvg = mean(Value, na.rm = TRUE), 
    AUC0InfPredAvg = mean(SimMetric, na.rm = TRUE),
    AUC0InfRatio = mean(SimMetric / Value, na.rm = TRUE)
  ) %>%
  summarise(
    AUC0InfRatio5th = 2^quantile(abs(log2(AUC0InfRatio)), c(0.05)),
    AUC0InfRatio25th = 2^quantile(abs(log2(AUC0InfRatio)), c(0.25)),
    AUC0InfRatio50th = 2^quantile(abs(log2(AUC0InfRatio)), c(0.50)),
    AUC0InfRatio75th = 2^quantile(abs(log2(AUC0InfRatio)), c(0.75)),
    AUC0InfRatio95th = 2^quantile(abs(log2(AUC0InfRatio)), c(0.95))
  )
```

If we look more in detail at the AUC 0-Inf 
(one dot is one studies and one compound i.e. one color can have multiple studies associated), 
we can see that most of the studies/compounds are within 10-fold of the observed value, with one 
compound being particularly underpredicted.

```{r}
metric_results_01 %>% filter(Metric == "AUC", `Metric_t1 [h]` == 0, `Metric_t2 [h]` == Inf) %>% 
  ggplot(aes(x = Value, y = SimMetric, color = molecule_pref_name, shape = Meal)) + 
  geom_point(show.legend = F) + 
  geom_abline(slope = 1, intercept = 0) + 
  labs(x = "Observed AUC0Inf", y = "Predicted AUC0Inf") + 
  scale_x_log10() + scale_y_log10() +
  geom_abline(slope = 1, intercept = log10(2), col = "grey50", linetype = "dashed") + 
  geom_abline(slope = 1, intercept = -log10(2), col = "grey50", linetype = "dashed") + 
  geom_abline(slope = 1, intercept = 1, col = "grey80", linetype = "dashed") + 
  geom_abline(slope = 1, intercept = -1, col = "grey80", linetype = "dashed") + 
  theme_bw()
```
Similarly we can also look at other metrics such as Cmax. 
The Cmax for all compounds tend to be overpredicted.

```{r echo=FALSE}
metric_results_01 %>% 
  filter(Metric == "Cmax", `Metric_t1 [h]` == 0) %>% 
  group_by(molecule_pref_name) %>%
  summarise(
    CmaxAvg = mean(Value, na.rm = TRUE), 
    CmaxPredAvg = mean(SimMetric, na.rm = TRUE),
    CmaxRatio = mean(SimMetric / Value, na.rm = TRUE)
  ) %>%
  ggplot() +
    geom_boxplot(aes(y = CmaxRatio)) + 
    scale_y_log10() + 
    theme_bw() + 
    theme(aspect.ratio = 5)
```

```{r include=FALSE}
tmp <- metric_results_01 %>% 
  filter(Metric == "Cmax", `Metric_t1 [h]` == 0) %>% 
  group_by(molecule_pref_name) %>%
  summarise(
    CmaxAvg = mean(Value, na.rm = TRUE), 
    CmaxPredAvg = mean(SimMetric, na.rm = TRUE),
    CmaxRatio = mean(SimMetric / Value, na.rm = TRUE)
  ) %>%
  summarise(
    CmaxRatio_2Fold = mean(CmaxRatio <= 2 & CmaxRatio >= 1/2, na.rm = TRUE) * 100,
    CmaxRatio_4Fold = mean(CmaxRatio <= 4 & CmaxRatio >= 1/4, na.rm = TRUE) * 100,
    CmaxRatio_10Fold = mean(CmaxRatio <= 10 & CmaxRatio >= 1/10, na.rm = TRUE) * 100,
  )
```

There are `r signif(tmp$CmaxRatio_2Fold,3` % of compounds with a predicted Cmax within 2-fold of the observed value, `r signif(tmp$CmaxRatio_4Fold, 3)`% within 4-fold and `r signif(tmp$CmaxRatio_10Fold,3)` % within 10-fold.

```{r}
metric_results_01 %>% 
  filter(Metric == "Cmax", `Metric_t1 [h]` == 0) %>% 
  group_by(molecule_pref_name) %>%
  summarise(
    CmaxAvg = mean(Value, na.rm = TRUE), 
    CmaxPredAvg = mean(SimMetric, na.rm = TRUE),
    CmaxRatio = mean(SimMetric / Value, na.rm = TRUE)
  ) %>%
  summarise(
    CmaxRatio5th = 2^quantile(abs(log2(CmaxRatio)), c(0.05)),
    CmaxRatio25th = 2^quantile(abs(log2(CmaxRatio)), c(0.25)),
    CmaxRatio50th = 2^quantile(abs(log2(CmaxRatio)), c(0.50)),
    CmaxRatio75th = 2^quantile(abs(log2(CmaxRatio)), c(0.75)),
    CmaxRatio95th = 2^quantile(abs(log2(CmaxRatio)), c(0.95))
  )
```

The comparison of Cmax can be slightly more difficult as the sample time points could affect the observed estimated value, 
and the full observed pk profiles are not available.

If we look more in detail at the Cmax 
(one dot is one studies and one compound i.e. one color can have multiple studies associated), 
we can see that most of the studies/compounds are within 10-fold of the observed value, with one 
compound being particularly underpredicted.


```{r}
metric_results_01 %>% filter(Metric == "Cmax", `Metric_t1 [h]` == 0) %>% 
  ggplot(aes(x = Value, y = SimMetric, color = molecule_pref_name, shape = Meal)) + 
  geom_point(show.legend = F) + 
  geom_abline(slope = 1, intercept = 0) + 
  labs(x = "Observed Cmax", y = "Predicted Cmax") + 
  scale_x_log10() + scale_y_log10() +
  geom_abline(slope = 1, intercept = log10(2), col = "grey50", linetype = "dashed") + 
  geom_abline(slope = 1, intercept = -log10(2), col = "grey50", linetype = "dashed") + 
  geom_abline(slope = 1, intercept = 1, col = "grey80", linetype = "dashed") + 
  geom_abline(slope = 1, intercept = -1, col = "grey80", linetype = "dashed") + 
  theme_bw()
```
Looking at AUC and Cmax ratio together:  
```{r}
metric_results_01 %>% 
  filter((Metric == "Cmax" & `Metric_t1 [h]` == 0 ) | (Metric == "AUC" & `Metric_t1 [h]` == 0 & `Metric_t2 [h]` == Inf)) %>% 
  tidyr::pivot_wider(
    id_cols = c(StudyID, ProtocolID, molecule_pref_name, canonical_smiles, Meal), 
    names_from = Metric, 
    values_from = c(Value, SimMetric), 
    values_fn = c("Value" = ~ mean(.x), "SimMetric" = ~ mean(.x))
  ) %>%
  mutate(AUC0Inf_Ratio = SimMetric_AUC / Value_AUC, 
         Cmax_Ratio = SimMetric_Cmax / Value_Cmax) %>%
  ggplot(aes(x = AUC0Inf_Ratio, y = Cmax_Ratio, color = molecule_pref_name, shape = Meal)) + 
  geom_point(show.legend = F) + 
  geom_abline(slope = 1, intercept = 0) + 
  labs(x = "AUC0Inf ratio", y = "Cmax ratio") + 
  scale_x_log10() + scale_y_log10() +
  geom_vline(xintercept = 2, col = "grey50", linetype = "dashed") + 
  geom_vline(xintercept = 0.5, col = "grey50", linetype = "dashed") + 
  geom_vline(xintercept = 10, col = "grey80", linetype = "dashed") + 
  geom_vline(xintercept = 0.1, col = "grey80", linetype = "dashed") + 
  geom_hline(yintercept = 2, col = "grey50", linetype = "dashed") + 
  geom_hline(yintercept = 0.5, col = "grey50", linetype = "dashed") + 
  geom_hline(yintercept = 10, col = "grey80", linetype = "dashed") + 
  geom_hline(yintercept = 0.1, col = "grey80", linetype = "dashed") + 
  theme_bw()
```

We can additional try to understand for which compounds the AUC and Cmax are less well predicted, 
to better understand the potential applicability domain of the pipeline even though
it will be mostly constrained by applicability domain of the used QSAR inputs as well as the PBPK 
model. 

Compounds with high lipophilicty or low lipophilicity values tends to be the 
most problematic and not well predicted. Similarly compounds that have a very low solubility are 
also poorly predicted
 
```{r}
full_join(
  x = metric_results_01 %>% 
    filter(Metric == "AUC" & `Metric_t1 [h]` == 0 & `Metric_t2 [h]` == Inf) %>% 
    mutate(AUC0Inf_Ratio = SimMetric / Value),
  y = InputCompounds, 
  by = c("molecule_pref_name", "canonical_smiles")
) %>% 
  ggplot(aes(x = logD, y = AUC0Inf_Ratio)) + 
  geom_point() + 
  geom_hline(yintercept = 2, col = "grey50", linetype = "dashed") + 
  geom_hline(yintercept = 0.5, col = "grey50", linetype = "dashed") + 
  geom_hline(yintercept = 10, col = "grey80", linetype = "dashed") + 
  geom_hline(yintercept = 0.1, col = "grey80", linetype = "dashed") + 
  scale_y_log10() + 
  labs(x = "Compound logD", y = "AUC0Inf ratio") + 
  theme_bw() 
```

```{r}
full_join(
  x = metric_results_01 %>% 
    filter(Metric == "AUC" & `Metric_t1 [h]` == 0 & `Metric_t2 [h]` == Inf) %>% 
    mutate(AUC0Inf_Ratio = SimMetric / Value),
  y = InputCompounds, 
  by = c("molecule_pref_name", "canonical_smiles")
) %>% 
  ggplot(aes(x = Fu, y = AUC0Inf_Ratio)) + 
  geom_point() + 
  geom_hline(yintercept = 2, col = "grey50", linetype = "dashed") + 
  geom_hline(yintercept = 0.5, col = "grey50", linetype = "dashed") + 
  geom_hline(yintercept = 10, col = "grey80", linetype = "dashed") + 
  geom_hline(yintercept = 0.1, col = "grey80", linetype = "dashed") + 
  scale_y_log10() + 
  labs(x = "Compound Fraction unbound", y = "AUC0Inf ratio") + 
  theme_bw() 
```

```{r}
full_join(
  x = metric_results_01 %>% 
    filter(Metric == "AUC" & `Metric_t1 [h]` == 0 & `Metric_t2 [h]` == Inf) %>% 
    mutate(AUC0Inf_Ratio = SimMetric / Value),
  y = InputCompounds, 
  by = c("molecule_pref_name", "canonical_smiles")
) %>% 
  ggplot(aes(x = logS, y = AUC0Inf_Ratio)) + 
  geom_point() + 
  geom_hline(yintercept = 2, col = "grey50", linetype = "dashed") + 
  geom_hline(yintercept = 0.5, col = "grey50", linetype = "dashed") + 
  geom_hline(yintercept = 10, col = "grey80", linetype = "dashed") + 
  geom_hline(yintercept = 0.1, col = "grey80", linetype = "dashed") + 
  scale_y_log10() + 
  labs(x = "Compound Solubility", y = "AUC0Inf ratio") + 
  theme_bw() 
```

We can test and compare with other scenarios, such as adding artificially high organ permeability (scenario 09)

```{r}
path <- system.file("extdata", "Chembl_Test_Pipeline", "metric_scenario_9.csv", package = "ESQhtpbpk")
metric_results_09 <- read.csv(path, stringsAsFactors = FALSE, check.names = FALSE)
``` 

Calculated and compare AUC 0 Inf ratio of both scenarios:

```{r echo=FALSE}
scenario_01_vs_09 <- rbind(
  metric_results_01 %>% 
    filter(Metric == "AUC", `Metric_t1 [h]` == 0, `Metric_t2 [h]` == Inf) %>% 
    group_by(molecule_pref_name) %>%
    summarise(
      AUC0InfAvg = mean(Value, na.rm = TRUE), 
      AUC0InfPredAvg = mean(SimMetric, na.rm = TRUE),
      AUC0InfRatio = mean(SimMetric / Value, na.rm = TRUE),
      Scenario = "01"
    ),
  metric_results_09 %>% 
    filter(Metric == "AUC", `Metric_t1 [h]` == 0, `Metric_t2 [h]` == Inf) %>% 
    group_by(molecule_pref_name) %>%
    summarise(
      AUC0InfAvg = mean(Value, na.rm = TRUE), 
      AUC0InfPredAvg = mean(SimMetric, na.rm = TRUE),
      AUC0InfRatio = mean(SimMetric / Value, na.rm = TRUE),
      Scenario = "09"
    )
)

full_join(
  x = scenario_01_vs_09 %>% pivot_wider(
    id_cols = c("molecule_pref_name"), 
    names_from = Scenario, 
    values_from = c(AUC0InfAvg, AUC0InfPredAvg, AUC0InfRatio)
  ),
  y = InputCompounds, 
  by = c("molecule_pref_name")
) %>% 
  ggplot(aes(x = AUC0InfRatio_01, y = AUC0InfRatio_09, color = logD)) + 
    geom_point() + 
    geom_abline(slope = 1, intercept = 0) + 
    labs(x = "AUC0Inf Ratio Scenario 01", y = "AUC0Inf Ratio Scenario 09") + 
    scale_x_log10() + scale_y_log10() +
  theme_bw()  

```
```{r echo=FALSE}
scenario_01_vs_09 <- rbind(
  metric_results_01 %>% 
    filter(Metric == "Cmax", `Metric_t1 [h]` == 0) %>% 
    group_by(molecule_pref_name) %>%
    summarise(
      CmaxAvg = mean(Value, na.rm = TRUE), 
      CmaxPredAvg = mean(SimMetric, na.rm = TRUE),
      CmaxRatio = mean(SimMetric / Value, na.rm = TRUE),
      Scenario = "01"
    ),
  metric_results_09 %>% 
    filter(Metric == "Cmax", `Metric_t1 [h]` == 0) %>% 
    group_by(molecule_pref_name) %>%
    summarise(
      CmaxAvg = mean(Value, na.rm = TRUE), 
      CmaxPredAvg = mean(SimMetric, na.rm = TRUE),
      CmaxRatio = mean(SimMetric / Value, na.rm = TRUE),
      Scenario = "09"
    )
)

full_join(
  x = scenario_01_vs_09 %>% pivot_wider(
    id_cols = c("molecule_pref_name"), 
    names_from = Scenario, 
    values_from = c(AUC0InfAvg, AUC0InfPredAvg, AUC0InfRatio)
  ),
  y = InputCompounds, 
  by = c("molecule_pref_name")
) %>%
  ggplot(aes(x = CmaxRatio_01, y = CmaxRatio_09, color = logD)) + 
    geom_point() + 
    geom_abline(slope = 1, intercept = 0) + 
    labs(x = "Observed Cmax Scenario 01", y = "Observed Cmax Scenario 09") + 
    scale_x_log10() + scale_y_log10() +
  theme_bw()  
``` 

This reduce the Cmax of all compounds and also the AUC of most compounds leading to overall
better results moslty by improving the Cmax twithout degrading to much the AUC, 
even though it is detrimental to some compounds.

```{r include=FALSE}
tmp <- metric_results_09 %>% 
  filter(Metric == "AUC", `Metric_t1 [h]` == 0, `Metric_t2 [h]` == Inf) %>% 
  group_by(molecule_pref_name) %>%
  summarise(
    AUC0InfAvg = mean(Value, na.rm = TRUE), 
    AUC0InfPredAvg = mean(SimMetric, na.rm = TRUE),
    AUC0InfRatio = mean(SimMetric / Value, na.rm = TRUE)
  ) %>%
  summarise(
    AUC0InfRatio_2Fold = mean(AUC0InfRatio <= 2 & AUC0InfRatio >= 1/2, na.rm = TRUE) * 100,
    AUC0InfRatio_4Fold = mean(AUC0InfRatio <= 4 & AUC0InfRatio >= 1/4, na.rm = TRUE) * 100,
    AUC0InfRatio_10Fold = mean(AUC0InfRatio <= 10 & AUC0InfRatio >= 1/10, na.rm = TRUE) * 100,
  )
```
There are `r signif(tmp$AUC0InfRatio_2Fold,3` % of compounds with a predicted AUC0Inf within 2-fold of the observed value, `r signif(tmp$AUC0InfRatio_4Fold, 3)`% within 4-fold and `r signif(tmp$AUC0InfRatio_10Fold,3)` % within 10-fold.

```{r}
metric_results_09 %>% 
  filter(Metric == "AUC", `Metric_t1 [h]` == 0, `Metric_t2 [h]` == Inf) %>% 
  group_by(molecule_pref_name) %>%
  summarise(
    AUC0InfAvg = mean(Value, na.rm = TRUE), 
    AUC0InfPredAvg = mean(SimMetric, na.rm = TRUE),
    AUC0InfRatio = mean(SimMetric / Value, na.rm = TRUE)
  ) %>%
  summarise(
    AUC0InfRatio5th = 2^quantile(abs(log2(AUC0InfRatio)), c(0.05)),
    AUC0InfRatio25th = 2^quantile(abs(log2(AUC0InfRatio)), c(0.25)),
    AUC0InfRatio50th = 2^quantile(abs(log2(AUC0InfRatio)), c(0.50)),
    AUC0InfRatio75th = 2^quantile(abs(log2(AUC0InfRatio)), c(0.75)),
    AUC0InfRatio95th = 2^quantile(abs(log2(AUC0InfRatio)), c(0.95))
  )
```

Similarly we can also look at other metrics such as Cmax. 
The Cmax for all compounds tend to be overpredicted.

```{r include=FALSE}
tmp <- metric_results_09 %>% 
  filter(Metric == "Cmax", `Metric_t1 [h]` == 0) %>% 
  group_by(molecule_pref_name) %>%
  summarise(
    CmaxAvg = mean(Value, na.rm = TRUE), 
    CmaxPredAvg = mean(SimMetric, na.rm = TRUE),
    CmaxRatio = mean(SimMetric / Value, na.rm = TRUE)
  ) %>%
  summarise(
    CmaxRatio_2Fold = mean(CmaxRatio <= 2 & CmaxRatio >= 1/2, na.rm = TRUE) * 100,
    CmaxRatio_4Fold = mean(CmaxRatio <= 4 & CmaxRatio >= 1/4, na.rm = TRUE) * 100,
    CmaxRatio_10Fold = mean(CmaxRatio <= 10 & CmaxRatio >= 1/10, na.rm = TRUE) * 100,
  )
```

There are `r signif(tmp$CmaxRatio_2Fold,3` % of compounds with a predicted Cmax within 2-fold of the observed value, `r signif(tmp$CmaxRatio_4Fold, 3)`% within 4-fold and `r signif(tmp$CmaxRatio_10Fold,3)` % within 10-fold.

```{r}
metric_results_09 %>% 
  filter(Metric == "Cmax", `Metric_t1 [h]` == 0) %>% 
  group_by(molecule_pref_name) %>%
  summarise(
    CmaxAvg = mean(Value, na.rm = TRUE), 
    CmaxPredAvg = mean(SimMetric, na.rm = TRUE),
    CmaxRatio = mean(SimMetric / Value, na.rm = TRUE)
  ) %>%
  summarise(
    CmaxRatio5th = 2^quantile(abs(log2(CmaxRatio)), c(0.05)),
    CmaxRatio25th = 2^quantile(abs(log2(CmaxRatio)), c(0.25)),
    CmaxRatio50th = 2^quantile(abs(log2(CmaxRatio)), c(0.50)),
    CmaxRatio75th = 2^quantile(abs(log2(CmaxRatio)), c(0.75)),
    CmaxRatio95th = 2^quantile(abs(log2(CmaxRatio)), c(0.95))
  )
```

